### 2022-06-30 report/paper.tex
Created latex report with initial problem description

\documentclass[12pt,a4paper]{article}
\usepackage{natbib}

\begin{document}
\title{Pulsar detection - Machine learning and pattern recognition exam}
\author{Alberto Baroso - s296520}
\date{2022-07-13}
\maketitle

\begin{abstract}
    The Pulsar dataset \cite{stw656} is a collection of 17.898 pulsar candidates of which 1.639 are real pulsars and 16,259 are spurious examples caused by RFI/noise.
    A pulsar is a neutron star that emits beams of electromagnetic radiation.
    As pulsars rotate, their emission beam sweeps across the sky, a periodically repeated pattern of broadband radio emission can be detected as this beam crosses our line of sight. 
    Thus pulsar search involves looking for periodic radio signals with large radio telescopes.
    The goal of this project is to use machine learning techniques to solve the binary classification problem of detecting pulsars from the Pulsar dataset.
\end{abstract}

\tableofcontents

\section{Problem analysis}

\subsection{Features}

Pulsar candidates are described by 8 continuous features, and a class label.

The first four are simple statistics obtained from the integrated pulse profile (folded profile). 
This is an array of continuous features that describe a longitude-resolved version of the signal that has been averaged in both time and frequency (see \cite{Lyon2016} for more details). 
The remaining four features are similarly obtained from the DM-SNR curve (again see \cite{Lyon2016} for more details). 
These are summarised below:

1. Mean of the integrated profile.
2. Standard deviation of the integrated profile.
3. Excess kurtosis of the integrated profile.
4. Skewness of the integrated profile.
5. Mean of the DM-SNR curve.
6. Standard deviation of the DM-SNR curve.
7. Excess kurtosis of the DM-SNR curve.
8. Skewness of the DM-SNR curve.
9. Class

The class labels used are 0 (negative) and 1 (positive).

\subsection{Feature distributions and ranges}

There are 8 features, that represent statistics extracted from radio signals collected by radio telescopes.
Classes are highly imbalanced.

\bibliographystyle{unsrt}
\bibliography{citations}

\end{document}

### 2022-07-02 report/paper.tex
Description of features and their distribution, plotted histograms and scatter plots

@@
 \usepackage{natbib}
+\usepackage{graphicx}
+\usepackage{float}
+\usepackage{multirow}
+\usepackage{subfig}
+\usepackage{alphalph}
+\renewcommand\thesubfigure{\alphalph{\value{subfigure}}}
 
@@
 \maketitle
+\clearpage
 
@@
 \end{abstract}
+\clearpage
 
 \tableofcontents
+\clearpage
 
@@
 
-Pulsar candidates are described by 8 continuous features, and a class label.
+Pulsar candidates are described by a class label and 8 continuous features extracted from radio signals collected by radio telescopes. \vspace{0.2cm}
 
@@
 The remaining four features are similarly obtained from the DM-SNR curve (again see \cite{Lyon2016} for more details).
-These are summarised below:
+\vspace{0.2cm}
+The features are listed below:
 
-1. Mean of the integrated profile.
-2. Standard deviation of the integrated profile.
-3. Excess kurtosis of the integrated profile.
-4. Skewness of the integrated profile.
-5. Mean of the DM-SNR curve.
-6. Standard deviation of the DM-SNR curve.
-7. Excess kurtosis of the DM-SNR curve.
-8. Skewness of the DM-SNR curve.
-9. Class
+\begin{enumerate}
+    \item Mean of the integrated profile.
+    \item  Standard deviation of the integrated profile.
+    \item Excess kurtosis of the integrated profile.
+    \item Skewness of the integrated profile.
+    \item Mean of the DM-SNR curve.
+    \item Standard deviation of the DM-SNR curve.
+    \item Excess kurtosis of the DM-SNR curve.
+    \item Skewness of the DM-SNR curve.
+    \item Class
+\end{enumerate}
 
-The class labels used are 0 (negative) and 1 (positive).
+The class labels used are 0 (non pulsar) and 1 (pulsar).
 
@@
 
-There are 8 features, that represent statistics extracted from radio signals collected by radio telescopes.
-Classes are highly imbalanced.
+\begin{figure}[H]
+    \begin{center}
+        \begin{tabular}{cc}
+            \subfloat[]{\includegraphics[width = 183pt, height = 120pt]{img/unprocessed_feature_distributions/mean_of_ip.png}}                &
+            \subfloat[]{\includegraphics[width = 183pt, height = 120pt]{img/unprocessed_feature_distributions/stdev_of_ip.png}}                 \\
+            \subfloat[]{\includegraphics[width = 183pt, height = 120pt]{img/unprocessed_feature_distributions/excess_kurtosis_of_ip.png}}     &
+            \subfloat[]{\includegraphics[width = 183pt, height = 120pt]{img/unprocessed_feature_distributions/skewness_of_ip.png}}              \\
+            \subfloat[]{\includegraphics[width = 183pt, height = 120pt]{img/unprocessed_feature_distributions/mean_of_dm-snr.png}}            &
+            \subfloat[]{\includegraphics[width = 183pt, height = 120pt]{img/unprocessed_feature_distributions/stdev_of_dm-snr.png}}             \\
+            \subfloat[]{\includegraphics[width = 183pt, height = 120pt]{img/unprocessed_feature_distributions/excess_kurtosis_of_dm-snr.png}} &
+            \subfloat[]{\includegraphics[width = 183pt, height = 120pt]{img/unprocessed_feature_distributions/skewness_of_dm-snr.png}}          \\
+        \end{tabular}
+    \end{center}
+\end{figure}
 
+\begin{minipage}{\linewidth}
+    \begin{tabular}{ |c|p{220pt}|p{34pt}|p{30pt}|p{36pt}|  }\hline
+
+           & Feature                                      & Mean   & Min   & Max     \\
+        \hline
+        a. & Mean of the integrated profile               & 110.86 & 6.18  & 186.02  \\
+        \hline
+        b. & Standard deviation of the integrated profile & 46.47  & 24.77 & 98.78   \\
+        \hline
+        c. & Excess kurtosis of the integrated profile    & 0.49   & -1.73 & 8.07    \\
+        \hline
+        d. & Skewness of the integrated profile           & 1.84   & -1.79 & 68.10   \\
+        \hline
+        e. & Mean of the DM-SNR curve                     & 12.67  & 0.21  & 209.30  \\
+        \hline
+        f. & Standard deviation of the DM-SNR curve       & 26.25  & 7.37  & 110.64  \\
+        \hline
+        g. & Excess kurtosis of the DM-SNR curve          & 8.33   & -2.81 & 34.54   \\
+        \hline
+        h. & Skewness of the DM-SNR curve                 & 105.41 & -1.98 & 1191.00 \\
+        \hline
+    \end{tabular}
+    \captionof{table}{Feature ranges}
+\end{minipage}
+\vspace{1cm}
+
+An initial analysis of the raw features of the training data shows that:
+\begin{itemize}
+    \item Features have mainly irregular distributions
+    \item Few features, the ones for non pulsars in figures (a), (b) and (g), present distributions similar to Gaussians
+    \item Looking at the graph (b) it's possible to deduce that the "Standard Deviation of the integrated profile" could be one of the less informative features as the histograms of the two classes share one of the largest overlapping areas compared to the other graphs.
+But since it also has a good amount of non overlapping areas it can still provide useful information for classification purposes.
+    \item Grphs (c) and (d) show that "Excess of kurtosis of integrated profile" and "Skewness of integrated profile" could be among the best features for sample discrimination, as they have little overlapping areas of the histograms.
+    \item Some features, mainly the ones in graphs (b), (d) and (h), are characterized by the presence of significantly large outliers.
+Similar claims can be made by looking at the ranges of minimum and maximum values for the corresponding features
+    \item We can expect classification algorithms to provide less than ideal results due to the presence of outliers.
+\end{itemize}
+
+% As a result, we further pre-process the data by "Gaussianizing" the features.
+
+\subsection{Feature pairs}
+
+\begin{figure}[H]
+    \begin{center}
+        \begin{tabular}{ccc}
+            \hspace*{-65pt}
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_ip-excess_kurtosis_dm-snr.png}} &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_ip-excess_kurtosis_ip.png}}     &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_ip-mean_dm-snr.png}}              \\
+            \hspace*{-65pt}
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_ip-skewness_dm-snr.png}}        &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_ip-skewness_ip.png}}            &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_ip-stdev_dm-snr.png}}             \\
+            \hspace*{-65pt}
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_ip-stdev_ip.png}}               &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_dm-snr-excess_kurtosis_ip.png}} &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_dm-snr-skewness_dm-snr.png}}      \\
+            \hspace*{-65pt}
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_dm-snr-skewness_ip.png}}        &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_dm-snr-stdev_dm-snr.png}}       &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/mean_dm-snr-stdev_ip.png}}             \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+\begin{figure}[H]
+    \begin{center}
+        \begin{tabular}{ccc}
+            \hspace*{-65pt}
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/excess_kurtosis_dm-snr-mean_dm-snr.png}}            &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/stdev_ip-stdev_dm-snr.png}}                     &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/stdev_ip-excess_kurtosis_dm-snr.png}}             \\
+            \hspace*{-65pt}
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/stdev_ip-excess_kurtosis_ip.png}}               &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/stdev_ip-skewness_ip.png}}                      &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/stdev_ip-skewness_dm-snr.png}}                    \\
+            \hspace*{-65pt}
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/stdev_dm-snr-skewness_ip.png}}                  &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/stdev_dm-snr-excess_kurtosis_ip.png}}           &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/stdev_dm-snr-skewness_dm-snr.png}}                \\
+            \hspace*{-65pt}
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/stdev_dm-snr-excess_kurtosis_dm-snr.png}}       &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/excess_kurtosis_ip-excess_kurtosis_dm-snr.png}} &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/excess_kurtosis_ip-skewness_dm-snr.png}}          \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+        \hspace*{-65pt}
+        \begin{tabular}{ccc}
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/excess_kurtosis_ip-skewness_ip.png}} &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/excess_kurtosis_dm-snr-excess_kurtosis_ip.png}} &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/excess_kurtosis_dm-snr-skewness_dm-snr.png}} \\
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/excess_kurtosis_dm-snr-skewness_ip.png}} &
+            \subfloat[]{\includegraphics[width = 160pt]{img/unprocessed_feature_pairs/skewness_ip-skewness_dm-snr.png}}  \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+% 
+% Classes are highly imbalanced.
+%
+
+Observing the feature pairs it's possible to note that there isn't one that allows to find a clear separation of the two classes. 
+\vspace*{0.5cm}
+The combination of features represented in graphs (n), (p), (ab), (af) and (ah) get close to identify regions of space where samples are well separated.
+\vspace*{0.5cm}\\
+The least useful pairs of features are represented in graphs (q), (s), (u) and (ac), where samples are almost entirely overlapping.
+
+% TODO: gaussianize features
+% TODO: feature correlation: compute pearson correlation heatmap
+
+\clearpage
 \bibliographystyle{unsrt}

### 2022-07-02 src/analytics/analytics.py

@@ -1,0 +1,24 @@
+import matplotlib.pyplot as plt
+import sys
+sys.path.append("./")
+from utils.data_load import load_training, load_features
+
+# Compute mean, min and max value of each feature
+def compute_analytics(features, samples):
+    print("Feature \t\t\t\t\t│  Mean \t│  Min \t\t│  Max")
+    print("─" * 90)
+    for i, feature in enumerate(features):
+        feature_name = feature + (" " * (40 - len(feature)))
+        mean_value = samples[i].mean()
+        min_value = samples[i].min()
+        max_value = samples[i].max()
+        print("{} \t│ {:.2f}  \t│ {:.2f}  \t│ {:.2f}".format(feature_name, mean_value, min_value, max_value))
+
+
+if __name__ == '__main__':
+
+    # Load data from file
+    samples, labels = load_training()
+    features = load_features()
+    
+    compute_analytics(features, samples)

### 2022-07-02 src/utils/data_load.py

@@ -1,0 +1,48 @@
+import numpy
+from utils.preprocessing import vcol
+
+
+def load_features():
+    return [
+        "Mean of the integrated profile",
+        "Standard deviation of the integrated profile",
+        "Excess kurtosis of the integrated profile",
+        "Skewness of the integrated profile",
+        "Mean of the DM-SNR curve", 
+        "Standard deviation of the DM-SNR curve",
+        "Excess kurtosis of the DM-SNR curve",
+        "Skewness of the DM-SNR curve"
+    ]
+
+
+def load_data(dataset):
+    samples = []
+    labels = []
+
+    # Read samples
+    with open(dataset, 'r') as file:
+        try:
+            for line in file:
+
+                attributes = line.split(",")
+
+                # Extract first 8 attributes as floats
+                sample = vcol(numpy.array([float(attr) for attr in attributes[0:8]]))
+
+                # Retrieve class label
+                label = int(attributes[-1].strip())
+
+                samples.append(sample)
+                labels.append(label)
+        except:
+            pass
+
+    return numpy.hstack(samples), numpy.array(labels)
+
+# Retrieve training samples
+def load_training():
+    return load_data('../data/Train.txt')
+
+# Retrieve test samples
+def load_test():
+    return load_data('../data/Test.txt')

### 2022-07-02 src/utils/preprocessing.py

@@ -1,0 +1,23 @@
+# Reshape a one-dimensional array to a column array
+def vcol(array):
+    return array.reshape((array.size, 1))
+
+# Reshape a column array to a row array
+def vrow(array):
+    return array.reshape((1, array.size))
+
+## DIMENSIONALITY  REDUCTION ##
+import numpy
+
+# Compute empirical mean
+def empirical_mean(array):
+    return vcol(array.mean(1))
+
+def covariance_matrix(D):
+    # Compute mean foreach column (feature)
+    mu = empirical_mean(D)
+    # Compute the 0-mean matrix (centered data)
+    DC = D - mu
+    # Compute covariance matrix
+    C = numpy.dot(DC, DC.T) / float(D.shape[1])
+    return C

### 2022-07-02 src/visualization/feature_plots.py

@@ -1,0 +1,49 @@
+import matplotlib.pyplot as plt
+import sys
+sys.path.append("./")
+from utils.data_load import load_training, load_features
+
+
+def plot_features_histograms(features, samples, labels):
+
+    # Get samples by class
+    non_pulsars = samples[:, labels == 0]
+    pulsars = samples[:, labels == 1]
+
+    # Plot histogram
+    for i in range(len(features)):
+        plt.hist(non_pulsars[i], bins=20, density=True, label="Non pulsars", alpha=0.4)
+        plt.hist(pulsars[i], bins=20, density=True, label="Pulsars", alpha=0.4)
+        plt.xlabel(features[i])
+        plt.legend()
+        plt.show()
+
+
+def plot_feature_pairs_sctterplots(features, samples, labels):
+
+    # Get samples by class
+    non_pulsars = samples[:, labels == 0]
+    pulsars = samples[:, labels == 1]
+
+    # Plot scatterplots for all pairs of different features
+    for i in range(len(features)):
+        for j in range(len(features)):
+            if i != j:
+
+                plt.plot(non_pulsars[i], non_pulsars[j], linestyle='', marker='.', markersize=10, label="Non pulsars")
+                plt.plot(pulsars[i], pulsars[j], linestyle='', marker='.', markersize=10, label="Pulsars")
+                plt.xlabel(features[i])
+                plt.ylabel(features[j])
+                
+                plt.legend()
+                plt.show()
+
+
+if __name__ == '__main__':
+
+    # Load data from file
+    samples, labels = load_training()
+    features = load_features()
+    
+    plot_features_histograms(features, samples, labels)
+    plot_feature_pairs_sctterplots(features, samples, labels)

### 2022-07-04 report/paper.tex
Added PCA, LDA and 4 MVG classifiers

@@
 Observing the feature pairs it's possible to note that there isn't one that allows to find a clear separation of the two classes.
-\vspace*{0.5cm}
+\vspace*{0.5cm}\\
 The combination of features represented in graphs (n), (p), (ab), (af) and (ah) get close to identify regions of space where samples are well separated.
@@
 The least useful pairs of features are represented in graphs (q), (s), (u) and (ac), where samples are almost entirely overlapping.
+\vspace*{0.5cm}\\
+The plots also show that some pair of features share similar distributions, implying that there's a correlation between them.
 
 % TODO: gaussianize features
-% TODO: feature correlation: compute pearson correlation heatmap
 
+\clearpage
+\subsection{Feature correlation}
+
+Computing the pearson correlation between all pairs of raw features and plotting it as a heatmap we obtain the following result:
+\begin{figure}[H]
+    \begin{center}
+        \subfloat[Pearson correlation heatmap]{\includegraphics[width = 320pt]{img/pearson_correlation/raw_heatmap.png}}
+    \end{center}
+\end{figure}
+The heatmap clearly shows that the following pairs of features are strongly correlated:
+\begin{enumerate}
+    \item (0) Mean of the integrated profile - (1) Standard deviation of the integrated profile.
+    \item (2) Excess kurtosis of the integrated profile - (3) Skewness of the integrated profile.
+    \item (4) Mean of the DM-SNR curve - (5) Standard deviation of the DM-SNR curve.
+    \item (6) Excess kurtosis of the DM-SNR curve - (7) Skewness of the DM-SNR curve.
+\end{enumerate}
+
+While the following pairs of features are weakly correlated, with a pearson coefficient close to 0.4:
+\begin{enumerate}
+    \item (2) Excess kurtosis of the integrated profile - (4) Mean of the DM-SNR curve
+    \item (2) Excess kurtosis of the integrated profile - (5) Standard deviation of the DM-SNR curve
+    \item (3) Skewness of the integrated profile - (4) Mean of the DM-SNR curve
+    \item (3) Skewness of the integrated profile - (5) Standard deviation of the DM-SNR curve
+\end{enumerate}
 \clearpage
@@
 
+% TODO: WHERE TO PUT THIS?
+Linear Discriminant Analysis (LDA) is deemed not useful as a preprocessing step for this project as LDA allows to find at most c - 1 discriminant directions, where c is the number of classes, thus we could find at most 1 direction.
+
 \end{document}

### 2022-07-04 src/analytics/analytics.py

@@ -1,24 +1,0 @@
-import matplotlib.pyplot as plt
-import sys
-sys.path.append("./")
-from utils.data_load import load_training, load_features
-
-# Compute mean, min and max value of each feature
-def compute_analytics(features, samples):
-    print("Feature \t\t\t\t\t│  Mean \t│  Min \t\t│  Max")
-    print("─" * 90)
-    for i, feature in enumerate(features):
-        feature_name = feature + (" " * (40 - len(feature)))
-        mean_value = samples[i].mean()
-        min_value = samples[i].min()
-        max_value = samples[i].max()
-        print("{} \t│ {:.2f}  \t│ {:.2f}  \t│ {:.2f}".format(feature_name, mean_value, min_value, max_value))
-
-
-if __name__ == '__main__':
-
-    # Load data from file
-    samples, labels = load_training()
-    features = load_features()
-    
-    compute_analytics(features, samples)

### 2022-07-04 src/data_processing/analytics.py

@@ -1,0 +1,65 @@
+import numpy
+import sys
+sys.path.append("./")
+from utils.data_load import load_training, load_features
+from utils.preprocessing import vcol    
+
+# Compute mean, min and max value of each feature
+def compute_analytics(features, samples):
+    print("Feature \t\t\t\t\t│  Mean \t│  Min \t\t│  Max")
+    print("─" * 90)
+    for i, feature in enumerate(features):
+        feature_name = feature + (" " * (40 - len(feature)))
+        mean_value = samples[i].mean()
+        min_value = samples[i].min()
+        max_value = samples[i].max()
+        print("{} \t│ {:.2f}  \t│ {:.2f}  \t│ {:.2f}".format(feature_name, mean_value, min_value, max_value))
+
+
+def pearson_correlation(samples):
+    return numpy.corrcoef(samples)
+
+# Compute empirical mean
+def empirical_mean(array):
+    return vcol(array.mean(1))
+
+
+## DIMENSIONALITY  REDUCTION ##
+def covariance_matrix(D):
+    # Compute mean foreach column (feature)
+    mu = empirical_mean(D)
+    # Compute the 0-mean matrix (centered data)
+    DC = D - mu
+    # Compute covariance matrix
+    C = numpy.dot(DC, DC.T) / float(D.shape[1])
+    return C
+
+
+# compute the log-densities for samples X
+def logpdf_GAU_ND(X, mu, C):
+    covariance_inverse = numpy.linalg.inv(C)
+    M = X.shape[0]
+    # the absolute value of the log-determinant is the second value
+    determinant_logarithm = numpy.linalg.slogdet(C)[1]
+    Y = numpy.array([])
+    for i in range(X.shape[1]):
+        x = X[:, i:i+1]
+        centered_sample = x - mu
+        log_density = - (M/2 * numpy.log(2 * numpy.pi)) - (1/2 * determinant_logarithm) - (
+            1/2 * numpy.dot(centered_sample.T, numpy.dot(covariance_inverse, centered_sample)))
+        Y = numpy.append(Y, log_density)
+    return Y.ravel()
+
+# Compute the density for a matrix of samples X
+def pdf(X, mu, C):
+    # Takes the exponent of the logarithm, so just the density not the log-density
+    return numpy.exp(logpdf_GAU_ND(X, mu, C))
+
+
+if __name__ == '__main__':
+
+    # Load data from file
+    samples, labels = load_training()
+    features = load_features()
+    
+    compute_analytics(features, samples)

### 2022-07-04 src/data_processing/dimensionality_reduction.py

@@ -1,0 +1,129 @@
+import numpy
+
+
+def lda(D, L, m):
+
+    from utils.analytics import covariance_matrix
+
+    def most_discriminant_egienvectors(C, m):
+        # Compute eigenvectors
+        s, U = numpy.linalg.eigh(C)
+        # s: eigenvalues sorted from smallest to largest
+        # U: eigenvectors (columns of U)
+
+        P = U[:, ::-1][:, 0:m]
+
+        # Since the covariance matrix is semi-definite positive,
+        # We can also get the sorted eigenvectors from the Singular Value Decomposition
+        U2, s2, Vh = numpy.linalg.svd(C)
+        # In this case, the singular values (which are equal to the eigenvalues) are sorted in descending order,
+        # the columns of U are the corresponding eigenvectors
+        P2 = U[:, 0:m]
+        return P
+
+
+    pulsar = D[:, L == 0]
+    non_pulsar = D[:, L == 1]
+
+    # Compute covariance matrix
+    C_setosa = covariance_matrix(pulsar)
+    C_versicolor = covariance_matrix(non_pulsar)
+
+    n_samples_setosa = pulsar.shape[1]
+    n_samples_versicolor = non_pulsar.shape[1]
+    N = n_samples_setosa + n_samples_versicolor
+    SW = ((C_setosa * n_samples_setosa) + (C_versicolor * n_samples_versicolor)) / N
+
+    # mu is the global mean of the dataset
+    mu = D.mean(1)
+    mu_setosa = vcol(pulsar.mean(axis=1) - mu)
+    mu_versicolor = vcol(non_pulsar.mean(axis=1) - mu)
+    SB = ((n_samples_setosa * numpy.dot(mu_setosa, mu_setosa.T)) +
+          (n_samples_versicolor * numpy.dot(mu_versicolor, mu_versicolor.T))) / N
+
+
+    ### SOLVE THE GENERALIZED EIGENVALUE PROBLEM ###
+    # m: at most n_of_classes - 1 discriminant directions
+
+    # Be careful, not numpy.linalg.eigh, because the numpy won't solve the generalized eigenvalue problem
+    s, U = scipy.linalg.eigh(SB, SW)
+    W = U[:, ::-1][:, 0:m]
+
+    # Solving the eigenvalue problem by joint diagonalization of SB and SW
+
[…]

### 2022-07-04 src/data_processing/error.py

@@ -1,0 +1,11 @@
+import sys
+sys.path.append("./")
+from utils.preprocessing import one_dimensional_array
+
+
+# Compute error rate as (#correctly_predicted / #total_samples)
+def error_rate(predicted, expected):
+    correct = sum(one_dimensional_array(predicted) == one_dimensional_array(expected))
+    accuracy = correct / expected.shape[1]
+    error = 1 - accuracy
+    return error / float(len(predicted))

### 2022-07-04 src/data_processing/preprocessing.py

@@ -1,0 +1,12 @@
+# Reshape a one-dimensional array to a column array
+def vcol(array):
+    return array.reshape((array.size, 1))
+
+
+# Reshape a column array to a row array
+def vrow(array):
+    return array.reshape((1, array.size))
+
+
+def one_dimensional_array(array):
+    return array.reshape((array.size,))

### 2022-07-04 src/models/multivariate_Gaussian_model.py

@@ -1,0 +1,75 @@
+import numpy
+import sys
+sys.path.append("./")
+from utils.preprocessing import vrow, vcol
+from utils.analytics import empirical_mean, covariance_matrix
+from scipy import special
+
+# MULTIVARIATE GAUSSIAN DISTRIBUTION
+# Compute log-density for a sample set of samples X
+#  M    : Size of the feature vector X
+#  X    : M × N matrix of samples x (It's a column vector)
+#  mu   : numpy array of shape (M, 1) (It's a column vector)
+#  C    : numpy array of shape (M, M) (It's a square matrix) representing the covariance matrix Σ
+# Returns Y, the vector of log-densities
+def logpdf_GAU_ND(X, mu, C):
+    covariance_inverse = numpy.linalg.inv(C)
+    M = X.shape[0]
+    # the absolute value of the log-determinant is the second value
+    determinant_logarithm = numpy.linalg.slogdet(C)[1]
+    Y = numpy.array([])
+    for i in range(X.shape[1]):
+        x = X[:, i:i+1]
+        centered_sample = x - mu
+        log_density = - (M/2 * numpy.log(2 * numpy.pi)) - (1/2 * determinant_logarithm) - (
+            1/2 * numpy.dot(centered_sample.T, numpy.dot(covariance_inverse, centered_sample)))
+        Y = numpy.append(Y, log_density)
+    return Y.ravel()
+
+
+# Compute the density for a matrix of samples X
+def pdf(X, mu, C):
+    # Takes the exponent of the logarithm, so just the density not the log-density
+    return numpy.exp(logpdf_GAU_ND(X, mu, C))
+
+## Multivariate Gaussian Classifier ##
+
+# the classifier assumes that samples of each class c ∈ {0, 1} can be modeled as samples of a multivariate
+# Gaussian distribution with class-dependent mean and covariance matrices, the solution is given by
+# the empirical mean and covariance matrix of each class
+def train_multivariate_gaussian_model(DTR, LTR, DTE, class_priors):
+    mu = []
+    C = []
+    for c in range(2):
+        data = DTR[:, LTR == c]
+        mu.append(empirical_mean(data))
+        C.append(covariance_matrix(data))
+
+    # The final goal is to compute class posterior probabilities P(c|x), we do that in 3 steps:
+
+    # 1st step: For each sample x, compute the likelihoods    
+    # S: Score matrix, S[i, j] = class conditional probability for sample j given class i
[…]

### 2022-07-04 src/models/naive_bayes_Gaussian_model.py

@@ -1,0 +1,62 @@
+import numpy
+import scipy
+import sys
+sys.path.append("./")
+from utils.preprocessing import vrow, vcol
+from utils.analytics import empirical_mean, covariance_matrix, logpdf_GAU_ND, pdf
+
+
+"""
+We now consider the Naive Bayes version of the classifier. 
+The Naive Bayes version of the MVG is simply a Gaussian classifier 
+where the covariance matrices are diagonal.
+"""
+
+## Naive Bayes Gaussian Classifier ##
+def train_naive_bayes_gaussian_model(DTR, LTR, DTE, class_priors):
+    mu = []
+    C = []
+    for c in range(2):
+        data = DTR[:, LTR == c]
+        # The ML solution for the mean parameters is the same as the MVG model
+        mu.append(empirical_mean(data))
+        # the ML solution for the covariance matrices is the diagonal of the ML solution for the MVG model
+        C.append(covariance_matrix(data) * numpy.identity(data.shape[0]))
+        """ 
+            Since the number of features is small, we can adapt the MVG code
+            by simply zeroing the out-of-diagonal elements of the MVG ML solution.
+            This can be done, for example, multiplying element-wise
+            the MVG ML solution with the identity matrix. The rest of the code remains unchanged. 
+            If we have large dimensional data, it may be advisable to implement ad-hoc 
+            functions to work directly with just
+            the diagonal of the covariance matrices (we won't do this in this course)
+        """
+
+    # The final goal is to compute class posterior probabilities P(c|x), we do that in 3 steps:
+
+    # 1st step: For each sample x, compute the likelihoods    
+    # S: Score matrix, S[i, j] = class conditional probability for sample j given class i
+    S = [pdf(DTE, mu[c], C[c]) for c in range(2)] 
+
+    # 2nd step: compute class posterior probabilities combining the score matrix with prior information 
+    # This requires multiplying each row of S by the prior probability of the corresponding class (1/2).
+    SJoint = S * vcol(numpy.array(class_priors)) # Joint likelihood for the samples and the corresponding class
+
+    # 3rd step: Compute class posterior probabilities
+    SMarginal = vrow(SJoint.sum(0)) # compute the marginal densities (i.e. for each sample sum the values for each class: SJoint[0, j] + SJoint[1, j] + SJoint[2, j])
[…]

### 2022-07-04 src/models/tied_covariance_Gaussian_model.py

@@ -1,0 +1,49 @@
+import numpy
+import scipy
+import sys
+sys.path.append("./")
+from utils.preprocessing import vrow, vcol
+from utils.analytics import empirical_mean, covariance_matrix, logpdf_GAU_ND, pdf
+
+## Tied covariance Gaussian Classifier ##
+def train_tied_covariance_gaussian_model(DTR, LTR, DTE, class_priors):
+    mu = []
+    C = numpy.zeros((DTR.shape[0], DTR.shape[0]))
+    total_samples = DTR.shape[1]
+    for c in range(2):
+        data = DTR[:, LTR == c]
+        # The ML solution for the mean parameters is the same as the MVG model
+        mu.append(empirical_mean(data))
+        # the ML solution for the covariance matrix is given by the empirical within-class covariance matrix
+        C += data.shape[1] * covariance_matrix(data)
+    C = (1 / total_samples) * C
+
+    # The final goal is to compute class posterior probabilities P(c|x), we do that in 3 steps:
+
+    # 1st step: For each sample x, compute the likelihoods    
+    # S: Score matrix, S[i, j] = class conditional probability for sample j given class i
+    S = [pdf(DTE, mu[c], C) for c in range(2)] 
+
+    # 2nd step: compute class posterior probabilities combining the score matrix with prior information 
+    # This requires multiplying each row of S by the prior probability of the corresponding class (1/2).
+    SJoint = S * vcol(numpy.array(class_priors)) # Joint likelihood for the samples and the corresponding class
+
+    # 3rd step: Compute class posterior probabilities
+    SMarginal = vrow(SJoint.sum(0)) # compute the marginal densities (i.e. for each sample sum the values for each class: SJoint[0, j] + SJoint[1, j] + SJoint[2, j])
+    SPost = SJoint / SMarginal # This broadcasts SMarginal so that each element of SJoint is divided by the corresponding element in SMarginal
+
+    # Get the class for each sample, the index of the row with highest SPost value
+    assigned_labels = numpy.argmax(SPost, axis=0)
+
+    log_density = S = [logpdf_GAU_ND(DTE, mu[c], C) for c in range(2)] 
+    log_prior_probabilities = vcol(numpy.log(class_priors))
+    logSJoint = log_density + log_prior_probabilities # matrix of joint densities
+    #  log-sum-exp trick  #
[…]

### 2022-07-04 src/models/tied_naive_bayes_Gaussian_model.py

@@ -1,0 +1,65 @@
+import numpy
+import scipy
+import sys
+sys.path.append("./")
+from utils.preprocessing import vrow, vcol
+from utils.analytics import empirical_mean, covariance_matrix, logpdf_GAU_ND, pdf
+
+"""
+We now consider the Naive Bayes version of the classifier. 
+The Naive Bayes version of the MVG is simply a Gaussian classifier 
+where the covariance matrices are diagonal.
+"""
+
+## Tied Naive Bayes Gaussian Classifier ##
+
+def train_tied_naive_bayes_gaussian_model(DTR, LTR, DTE, class_priors):
+    mu = []
+    C = numpy.zeros((DTR.shape[0], DTR.shape[0]))
+
+    for c in range(2):
+        data = DTR[:, LTR == c]
+        # The ML solution for the mean parameters is the same as the MVG model
+        mu.append(empirical_mean(data))
+        # the ML solution for the covariance matrices is the diagonal of the ML solution for the MVG model
+        C += data.shape[1] * covariance_matrix(data)
+    C = (1 / DTR.shape[1]) * C
+    C = C * numpy.identity(C.shape[0])
+    """ 
+        Since the number of features is small, we can adapt the MVG code
+        by simply zeroing the out-of-diagonal elements of the MVG ML solution.
+        This can be done, for example, multiplying element-wise
+        the MVG ML solution with the identity matrix. The rest of the code remains unchanged. 
+        If we have large dimensional data, it may be advisable to implement ad-hoc 
+        functions to work directly with just
+        the diagonal of the covariance matrices (we won't do this in this course)
+    """
+
+    # The final goal is to compute class posterior probabilities P(c|x), we do that in 3 steps:
+
+    # 1st step: For each sample x, compute the likelihoods    
+    # S: Score matrix, S[i, j] = class conditional probability for sample j given class i
+    S = [pdf(DTE, mu[c], C) for c in range(2)] 
+
+    # 2nd step: compute class posterior probabilities combining the score matrix with prior information 
+    # This requires multiplying each row of S by the prior probability of the corresponding class (1/2).
+    SJoint = S * vcol(numpy.array(class_priors)) # Joint likelihood for the samples and the corresponding class
+
+    # 3rd step: Compute class posterior probabilities
[…]

### 2022-07-04 src/models/train.py

@@ -1,0 +1,75 @@
+import numpy
+import sys
+sys.path.append("./")
+from utils.data_load import load_training, load_test
+from tied_naive_bayes_Gaussian_model import train_tied_naive_bayes_gaussian_model
+from tied_covariance_Gaussian_model import train_tied_covariance_gaussian_model
+from multivariate_Gaussian_model import train_multivariate_gaussian_model
+from naive_bayes_Gaussian_model import train_naive_bayes_gaussian_model
+from utils.error import error_rate
+from utils.preprocessing import lda, pca
+
+def train_models(training_samples, training_labels, test_samples, test_labels, class_priors):
+    
+    # Train models & Predict labels
+    predictions_mvg = train_multivariate_gaussian_model(
+        training_samples, training_labels, test_samples, class_priors)
+    predictions_naive_bayes = train_naive_bayes_gaussian_model(
+        training_samples, training_labels, test_samples, class_priors)
+    predictions_tied_covariance = train_tied_covariance_gaussian_model(
+        training_samples, training_labels, test_samples, class_priors)
+    predictions_tied_naive_bayes = train_tied_naive_bayes_gaussian_model(
+        training_samples, training_labels, test_samples, class_priors)
+
+    predictions_mvg = numpy.array([predictions_mvg])
+    predictions_naive_bayes = numpy.array([predictions_naive_bayes])
+    predictions_tied_covariance = numpy.array([predictions_tied_covariance])
+    predictions_tied_naive_bayes = numpy.array([predictions_tied_naive_bayes])
+    test_labels = numpy.array([test_labels])
+
+    # Compute error rates
+    mvg_error = error_rate(predictions_mvg, test_labels)
+    naive_bayes_error = error_rate(predictions_naive_bayes, test_labels)
+    mvg_tied_cov_error = error_rate(predictions_tied_covariance, test_labels)
+    tied_naive_bayes_error = error_rate(predictions_tied_naive_bayes, test_labels)
+    return mvg_error, naive_bayes_error, mvg_tied_cov_error, tied_naive_bayes_error
+
+
+if __name__ == '__main__':
+
+    training_samples, training_labels = load_training()
+    test_samples, test_labels = load_test()
+    class_priors_vec = [[0.5, 0.5], [0.2, 0.8], [0.8, 0.2]]
+    
+    for class_priors in class_priors_vec:
+    
+        # Raw features
[…]

### 2022-07-04 src/utils/preprocessing.py

@@ -1,23 +1,0 @@
-# Reshape a one-dimensional array to a column array
-def vcol(array):
-    return array.reshape((array.size, 1))
-
-# Reshape a column array to a row array
-def vrow(array):
-    return array.reshape((1, array.size))
-
-## DIMENSIONALITY  REDUCTION ##
-import numpy
-
-# Compute empirical mean
-def empirical_mean(array):
-    return vcol(array.mean(1))
-
-def covariance_matrix(D):
-    # Compute mean foreach column (feature)
-    mu = empirical_mean(D)
-    # Compute the 0-mean matrix (centered data)
-    DC = D - mu
-    # Compute covariance matrix
-    C = numpy.dot(DC, DC.T) / float(D.shape[1])
-    return C

### 2022-07-04 src/visualization/feature_plots.py

@@ -2,6 +2,7 @@ import matplotlib.pyplot as plt
 import sys
 sys.path.append("./")
 from utils.data_load import load_training, load_features
+from utils.analytics import pearson_correlation
 
 
 def plot_features_histograms(features, samples, labels):
@@ -39,11 +40,19 @@ def plot_feature_pairs_sctterplots(features, samples, labels):
                 plt.show()
 
 
+def plot_correlation_heatmap(samples):
+    pearson_matrix = pearson_correlation(samples)
+    plt.imshow(pearson_matrix, cmap='gist_yarg', vmin=-1, vmax=1)
+    plt.show()
+    
+
+
 if __name__ == '__main__':
 
     # Load data from file
     samples, labels = load_training()
     features = load_features()
     
-    plot_features_histograms(features, samples, labels)
-    plot_feature_pairs_sctterplots(features, samples, labels)
+    plot_correlation_heatmap(samples)
+    # plot_features_histograms(features, samples, labels)
+    # plot_feature_pairs_sctterplots(features, samples, labels)

### 2022-07-05 src/data_processing/analytics.py
Implemented gaussianization of features and logistic regression model

@@ -1,8 +1,8 @@
 import numpy
 import sys
 sys.path.append("./")
-from utils.data_load import load_training, load_features
-from utils.preprocessing import vcol    
+from data_processing.data_load import load_training, load_features
+from data_processing.utils import vcol    
 
 # Compute mean, min and max value of each feature
 def compute_analytics(features, samples):

### 2022-07-05 src/data_processing/data_load.py

@@ -1,5 +1,7 @@
 import numpy
-from utils.preprocessing import vcol
+import sys
+sys.path.append("./")
+from data_processing.utils import vcol
 
 
 def load_features():

### 2022-07-05 src/data_processing/dimensionality_reduction.py

@@ -1,9 +1,13 @@
 import numpy
+import scipy
+import sys
+sys.path.append("./")
+from data_processing.analytics import covariance_matrix
+from data_processing.utils import vcol
 
 
 def lda(D, L, m):
 
-    from utils.analytics import covariance_matrix
 
     def most_discriminant_egienvectors(C, m):
         # Compute eigenvectors
@@ -108,7 +112,6 @@ def pca(D, m, labels):
 
     # Compute covariance matrix
     C = numpy.dot(DC, DC.T) / float(D.shape[1])
-    from utils.analytics import covariance_matrix
     C2 = covariance_matrix(DC)
 
     # Retrieve The m leading eigenvectors from U (Principal components)

### 2022-07-05 src/data_processing/error.py

@@ -1,6 +1,6 @@
 import sys
 sys.path.append("./")
-from utils.preprocessing import one_dimensional_array
+from data_processing.utils import one_dimensional_array
 
 
 # Compute error rate as (#correctly_predicted / #total_samples)

### 2022-07-05 src/data_processing/gaussianization.py

@@ -1,0 +1,57 @@
+from scipy.stats import norm
+import numpy
+
+### GAUSSIANIZATION ###
+
+# Compute percentage of samples smaller than x
+def rank_samples(feature, training_samples):
+    comparison_feature = feature if training_samples is None else training_samples
+    ranks = []
+    for x in feature:
+        less_than_x = 0
+        for x_i in comparison_feature: # x_i is the value of the considered feature for the i-th training sample
+            if x_i < x:
+                less_than_x += 1
+        rank = (less_than_x + 1) / (feature.shape[0] + 2)
+        ranks.append(rank)
+    # +2 because we assume the existance of a feature smaller than all the others and a feature larger than all the others)
+    return ranks
+
+# Compute percentile (percent poin function) of the rank
+def rank_percentile(rank):
+    return norm.ppf(rank)
+
+def gaussianize_feature(feature, training_samples=None):
+    return numpy.array(rank_percentile(rank_samples(feature, training_samples)))
+    
+def gaussianize_training_samples(DTR):
+    return numpy.array([gaussianize_feature(x) for x in DTR])
+
+def gaussianize_test_features(DTE, DTR):
+    # Compute gaussianization by raking test samples against training samples
+    result = numpy.array([])
+    for i in range(len(DTE)):
+        result = numpy.append(result, gaussianize_feature(DTE[i], DTR[i]))
+    return result
+    
+if __name__ == '__main__': 
+    
+    import sys
+    sys.path.append("./")
+    from data_processing.data_load import load_training, load_features, load_test
+    from visualization.feature_plots import plot_feature_pairs_sctterplots
+    import matplotlib.pyplot as plt
+    
+    features = load_features()
+    training_samples, training_labels = load_training()
+    for i in range(len(features)):
+    
+        gaussianized_training_samples = gaussianize_feature(training_samples[i])
+        non_pulsars = gaussianized_training_samples[training_labels == 0]
+        pulsars = gaussianized_training_samples[training_labels == 1]
+    
+        plt.hist(non_pulsars, bins=50, density=True, label="Non pulsars", alpha=0.4)
+        plt.hist(pulsars, bins=50, density=True, label="Pulsars", alpha=0.4)
+        plt.xlabel(features[i])
+        plt.legend()
+        plt.show()

### 2022-07-05 src/models/logistic_regression_model.py

@@ -1,0 +1,49 @@
+import numpy
+import scipy.optimize
+import sys
+sys.path.append("./")
+from data_processing.utils import vcol 
+from data_processing.error import error_rate 
+
+
+# LOGISTIC REGRESSION
+
+# DTR training data
+# LTR training labels
+# DTE test data
+# λ is a hyper-parameter
+def train_logistic_regression_model(DTR, LTR, DTE, λ):
+
+    # v is and array with shape (D+1,)
+    # D is the dimensionality of the feature space
+    # λ is a hyper-parameter
+    def logreg_obj(v, DTR, LTR, λ):
+        M = DTR.shape[0]
+        N = DTR.shape[1]
+        w, b = vcol(v[0:M]), v[-1]
+        cross_entropy = 0
+        for i in range(N):
+            c_i = LTR[i]
+            x_i = DTR[:, i:i+1]
+            z_i = 2.0 * c_i - 1.0
+            s = numpy.dot(w.T, x_i) + b
+            # numpy.logaddexp used to sum e^a + e^b without numerical issues, == (1 + e^(-z_i*...))
+            cross_entropy += numpy.logaddexp(0, - s * z_i)
+        regularization_term = (λ / 2) * (numpy.linalg.norm(w) ** 2)
+        return regularization_term + 1/N * cross_entropy
+        
+
+    # numerical solver to minimize the logreg_obj function
+    x, f, d = scipy.optimize.fmin_l_bfgs_b(func=logreg_obj, iprint=1, x0=numpy.zeros(DTR.shape[0] + 1), args=(DTR, LTR, λ), approx_grad=True)
+
+    # Compute scores as posterior log-likelihood ratios:
+    w = x[0:DTR.shape[0]]
+    b = x[-1]
+    predicted = []
+    for i in range(DTE.shape[1]):
+        x_t = DTE[:, i:i+1]
+        score_x_t = numpy.dot(w.T, x_t) + b
+        label_x_t = 1 if score_x_t >= 0 else 0
+        predicted.append(label_x_t)
+        
+    return predicted

### 2022-07-05 src/models/multivariate_Gaussian_model.py

@@ -1,8 +1,8 @@
 import numpy
 import sys
 sys.path.append("./")
-from utils.preprocessing import vrow, vcol
-from utils.analytics import empirical_mean, covariance_matrix
+from data_processing.utils import vrow, vcol
+from data_processing.analytics import empirical_mean, covariance_matrix
 from scipy import special
 
 # MULTIVARIATE GAUSSIAN DISTRIBUTION

### 2022-07-05 src/models/naive_bayes_Gaussian_model.py

@@ -2,8 +2,8 @@ import numpy
 import scipy
 import sys
 sys.path.append("./")
-from utils.preprocessing import vrow, vcol
-from utils.analytics import empirical_mean, covariance_matrix, logpdf_GAU_ND, pdf
+from data_processing.utils import vrow, vcol
+from data_processing.analytics import empirical_mean, covariance_matrix, logpdf_GAU_ND, pdf
 
 
 """

### 2022-07-05 src/models/tied_covariance_Gaussian_model.py

@@ -2,8 +2,8 @@ import numpy
 import scipy
 import sys
 sys.path.append("./")
-from utils.preprocessing import vrow, vcol
-from utils.analytics import empirical_mean, covariance_matrix, logpdf_GAU_ND, pdf
+from data_processing.utils import vrow, vcol
+from data_processing.analytics import empirical_mean, covariance_matrix, logpdf_GAU_ND, pdf
 
 ## Tied covariance Gaussian Classifier ##
 def train_tied_covariance_gaussian_model(DTR, LTR, DTE, class_priors):

### 2022-07-05 src/models/tied_naive_bayes_Gaussian_model.py

@@ -2,8 +2,8 @@ import numpy
 import scipy
 import sys
 sys.path.append("./")
-from utils.preprocessing import vrow, vcol
-from utils.analytics import empirical_mean, covariance_matrix, logpdf_GAU_ND, pdf
+from data_processing.utils import vrow, vcol
+from data_processing.analytics import empirical_mean, covariance_matrix, logpdf_GAU_ND, pdf
 
 """
 We now consider the Naive Bayes version of the classifier. 

### 2022-07-05 src/models/train.py

@@ -1,16 +1,18 @@
 import numpy
 import sys
 sys.path.append("./")
-from utils.data_load import load_training, load_test
-from tied_naive_bayes_Gaussian_model import train_tied_naive_bayes_gaussian_model
-from tied_covariance_Gaussian_model import train_tied_covariance_gaussian_model
-from multivariate_Gaussian_model import train_multivariate_gaussian_model
+from data_processing.dimensionality_reduction import lda, pca
+from data_processing.error import error_rate
 from naive_bayes_Gaussian_model import train_naive_bayes_gaussian_model
-from utils.error import error_rate
-from utils.preprocessing import lda, pca
+from multivariate_Gaussian_model import train_multivariate_gaussian_model
+from tied_covariance_Gaussian_model import train_tied_covariance_gaussian_model
+from tied_naive_bayes_Gaussian_model import train_tied_naive_bayes_gaussian_model
+from logistic_regression_model import train_logistic_regression_model
+from data_processing.data_load import load_training, load_test
 
-def train_models(training_samples, training_labels, test_samples, test_labels, class_priors):
-    
+
+def train_mvg_models(training_samples, training_labels, test_samples, test_labels, class_priors):
+
     # Train models & Predict labels
     predictions_mvg = train_multivariate_gaussian_model(
         training_samples, training_labels, test_samples, class_priors)
@@ -20,19 +22,39 @@ def train_models(training_samples, training_labels, test_samples, test_labels, c
         training_samples, training_labels, test_samples, class_priors)
     predictions_tied_naive_bayes = train_tied_naive_bayes_gaussian_model(
         training_samples, training_labels, test_samples, class_priors)
-
+    
     predictions_mvg = numpy.array([predictions_mvg])
     predictions_naive_bayes = numpy.array([predictions_naive_bayes])
     predictions_tied_covariance = numpy.array([predictions_tied_covariance])
     predictions_tied_naive_bayes = numpy.array([predictions_tied_naive_bayes])
+
     test_labels = numpy.array([test_labels])
 
     # Compute error rates
     mvg_error = error_rate(predictions_mvg, test_labels)
     naive_bayes_error = error_rate(predictions_naive_bayes, test_labels)
     mvg_tied_cov_error = error_rate(predictions_tied_covariance, test_labels)
[…]

### 2022-07-05 src/visualization/feature_plots.py

@@ -1,8 +1,8 @@
 import matplotlib.pyplot as plt
 import sys
 sys.path.append("./")
-from utils.data_load import load_training, load_features
-from utils.analytics import pearson_correlation
+from data_processing.data_load import load_training, load_features
+from data_processing.analytics import pearson_correlation
 
 
 def plot_features_histograms(features, samples, labels):
@@ -54,5 +54,5 @@ if __name__ == '__main__':
     features = load_features()
     
     plot_correlation_heatmap(samples)
-    # plot_features_histograms(features, samples, labels)
-    # plot_feature_pairs_sctterplots(features, samples, labels)
+    plot_features_histograms(features, samples, labels)
+    plot_feature_pairs_sctterplots(features, samples, labels)

### 2022-07-06 report/paper.tex
Feature preprocessing saved to file

@@
 
-% TODO: gaussianize features
-
 \clearpage
@@
     \begin{center}
-        \subfloat[Pearson correlation heatmap]{\includegraphics[width = 320pt]{img/pearson_correlation/raw_heatmap.png}}
+        \subfloat[Pearson correlation heatmap RAW]{\includegraphics[width = 320pt]{img/pearson_correlation/raw_heatmap.png}}
     \end{center}
@@
 \end{enumerate}
+
+The results from the correlation analysis suggest that models might benefit from applying a dimensionality reduction technique such as PCA to reduce the feature space to 4 uncorrelated features.
+\begin{figure}[H]
+    \begin{center}
+        \subfloat[Pearson correlation heatmap PCA]{\includegraphics[width = 320pt]{img/pearson_correlation/pca_heatmap.png}}
+    \end{center}
+\end{figure}
+Recomputing the pearson correlation and plotting it as a heatmap it's possible to see that the remaining 4 features are now uncorrelated.
+
 \clearpage
@@
 
+We expect the MVG with naive bayes assumption model to perform worse that the others on the raw feature because we have seen from the correlation heatmap that some features are strongly correlated and by definition the naive bayes model assumes that the features are independent.
+One other reason to believe that this model will have poor perform is that it is usually employed 
+%TODO: CHECK THIS and improve it
+when the samples are few and dimensions are a lot
+
 \end{document}

### 2022-07-06 src/data_processing/dimensionality_reduction.py

@@ -30,20 +30,20 @@ def lda(D, L, m):
     non_pulsar = D[:, L == 1]
 
     # Compute covariance matrix
-    C_setosa = covariance_matrix(pulsar)
-    C_versicolor = covariance_matrix(non_pulsar)
+    C_pulsar = covariance_matrix(pulsar)
+    C_non_pulsar = covariance_matrix(non_pulsar)
 
-    n_samples_setosa = pulsar.shape[1]
-    n_samples_versicolor = non_pulsar.shape[1]
-    N = n_samples_setosa + n_samples_versicolor
-    SW = ((C_setosa * n_samples_setosa) + (C_versicolor * n_samples_versicolor)) / N
+    n_samples_pulsar = pulsar.shape[1]
+    n_samples_non_pulsar = non_pulsar.shape[1]
+    N = n_samples_pulsar + n_samples_non_pulsar
+    SW = ((C_pulsar * n_samples_pulsar) + (C_non_pulsar * n_samples_non_pulsar)) / N
 
     # mu is the global mean of the dataset
     mu = D.mean(1)
-    mu_setosa = vcol(pulsar.mean(axis=1) - mu)
-    mu_versicolor = vcol(non_pulsar.mean(axis=1) - mu)
-    SB = ((n_samples_setosa * numpy.dot(mu_setosa, mu_setosa.T)) +
-          (n_samples_versicolor * numpy.dot(mu_versicolor, mu_versicolor.T))) / N
+    mu_pulsar = vcol(pulsar.mean(axis=1) - mu)
+    mu_non_pulsar = vcol(non_pulsar.mean(axis=1) - mu)
+    SB = ((n_samples_pulsar * numpy.dot(mu_pulsar, mu_pulsar.T)) +
+          (n_samples_non_pulsar * numpy.dot(mu_non_pulsar, mu_non_pulsar.T))) / N
 
 
     ### SOLVE THE GENERALIZED EIGENVALUE PROBLEM ###
@@ -112,21 +112,20 @@ def pca(D, m, labels):
 
     # Compute covariance matrix
     C = numpy.dot(DC, DC.T) / float(D.shape[1])
-    C2 = covariance_matrix(DC)
 
     # Retrieve The m leading eigenvectors from U (Principal components)
     # (We reverse the order of the columns of U so that the leading eigenvectors are in the first m columns):
     directions = pca_projection_matrix(C, m)
 
-    # Scatterplot
-    # pulsar = D[:, labels == 1]
[…]

### 2022-07-06 src/data_processing/gaussianization.py

@@ -27,31 +27,30 @@ def gaussianize_feature(feature, training_samples=None):
 def gaussianize_training_samples(DTR):
     return numpy.array([gaussianize_feature(x) for x in DTR])
 
-def gaussianize_test_features(DTE, DTR):
+def gaussianize_test_samples(DTE, DTR):
     # Compute gaussianization by raking test samples against training samples
-    result = numpy.array([])
+    result = []
     for i in range(len(DTE)):
-        result = numpy.append(result, gaussianize_feature(DTE[i], DTR[i]))
-    return result
+        result.append(gaussianize_feature(DTE[i], DTR[i]))
+    return numpy.vstack(result)
     
 if __name__ == '__main__': 
     
     import sys
     sys.path.append("./")
     from data_processing.data_load import load_training, load_features, load_test
-    from visualization.feature_plots import plot_feature_pairs_sctterplots
+    from visualization.feature_plots import plot_feature_histogram
     import matplotlib.pyplot as plt
     
     features = load_features()
     training_samples, training_labels = load_training()
+    test_samples, test_labels = load_test()
+    
+    # GAUSSIANIZATION OF SAMPLES #
+    gaussianized_training_samples = gaussianize_training_samples(training_samples)
+    gaussianized_test_samples = gaussianize_test_samples(test_samples, training_samples)
+    
     for i in range(len(features)):
-    
-        gaussianized_training_samples = gaussianize_feature(training_samples[i])
-        non_pulsars = gaussianized_training_samples[training_labels == 0]
-        pulsars = gaussianized_training_samples[training_labels == 1]
-    
-        plt.hist(non_pulsars, bins=50, density=True, label="Non pulsars", alpha=0.4)
-        plt.hist(pulsars, bins=50, density=True, label="Pulsars", alpha=0.4)
-        plt.xlabel(features[i])
-        plt.legend()
-        plt.show()
+        plot_feature_histogram(features[i], gaussianized_training_samples[i], training_labels)
+        plot_feature_histogram(features[i], gaussianized_test_samples[i], test_labels)
+        

### 2022-07-06 src/models/train.py

@@ -1,5 +1,6 @@
 import numpy
 import sys
+from os.path import exists
 sys.path.append("./")
 from data_processing.dimensionality_reduction import lda, pca
 from data_processing.error import error_rate
@@ -9,6 +10,7 @@ from tied_covariance_Gaussian_model import train_tied_covariance_gaussian_model
 from tied_naive_bayes_Gaussian_model import train_tied_naive_bayes_gaussian_model
 from logistic_regression_model import train_logistic_regression_model
 from data_processing.data_load import load_training, load_test
+from data_processing.gaussianization import gaussianize_test_samples,  gaussianize_training_samples
 
 
 def train_mvg_models(training_samples, training_labels, test_samples, test_labels, class_priors):
@@ -59,58 +61,133 @@ def train_log_reg_model(training_samples, training_labels, test_samples, test_la
 
 if __name__ == '__main__':
 
+    # Data load & preprocessing #
+    
     training_samples, training_labels = load_training()
     test_samples, test_labels = load_test()
-    class_priors_vec = [[0.5, 0.5], [0.2, 0.8], [0.8, 0.2]]
-    log_reg_λ_vec = [0, 0.0001, 0.01, 0.1]
-
-    for class_priors in class_priors_vec:
-
-        ################
-        # Raw features #
-        ################
-
-        (
-            mvg_error, 
-            naive_bayes_error, 
-            mvg_tied_cov_error, 
-            tied_naive_bayes_error,
-        ) = train_mvg_models(training_samples, training_labels, test_samples, test_labels, class_priors)
-
-        # print("# MVG ERROR RATES, priors: {} #".format(class_priors))
-        # print("Multivariate Gaussian Model: ", mvg_error)
-        # print("Naive bayes Model: ", naive_bayes_error)
-        # print("Tied covariance Model: ", mvg_tied_cov_error)
-        # print("Tied naive bayes Model: ", tied_naive_bayes_error)
-
-        # # LDA
-        # lda_training_samples = lda(training_samples, training_labels, 1)
-        # lda_test_samples = lda(test_samples, test_labels, 1)
-        # mvg_error, naive_bayes_error, mvg_tied_cov_error, tied_naive_bayes_error = train_models(lda_training_samples, training_labels, lda_test_samples, test_labels, class_priors)
-
-        # print("# LDA ERROR RATES, priors: {} #".format(class_priors))
[…]

### 2022-07-06 src/visualization/feature_plots.py

@@ -3,8 +3,9 @@ import sys
 sys.path.append("./")
 from data_processing.data_load import load_training, load_features
 from data_processing.analytics import pearson_correlation
+from data_processing.dimensionality_reduction import pca
 
-
+# Plot histogram for all features
 def plot_features_histograms(features, samples, labels):
 
     # Get samples by class
@@ -19,6 +20,19 @@ def plot_features_histograms(features, samples, labels):
         plt.legend()
         plt.show()
 
+# Plot histogram for only one feature
+def plot_feature_histogram(feature, samples, labels):
+
+    # Get samples by class
+    non_pulsars = samples[labels == 0]
+    pulsars = samples[labels == 1]
+
+    # Plot histogram
+    plt.hist(non_pulsars, bins=50, density=True, label="Non pulsars", alpha=0.4)
+    plt.hist(pulsars, bins=50, density=True, label="Pulsars", alpha=0.4)
+    plt.xlabel(feature)
+    plt.legend()
+    plt.show()
 
 def plot_feature_pairs_sctterplots(features, samples, labels):
 
@@ -53,6 +67,9 @@ if __name__ == '__main__':
     samples, labels = load_training()
     features = load_features()
     
+    pca_samples = pca(samples, 4, labels)
+    
     plot_correlation_heatmap(samples)
-    plot_features_histograms(features, samples, labels)
-    plot_feature_pairs_sctterplots(features, samples, labels)
+    plot_correlation_heatmap(pca_samples)
+    # plot_features_histograms(features, samples, labels)
+    # plot_feature_pairs_sctterplots(features, samples, labels)

### 2023-03-11 report/paper.tex
Completed report

--- 2022-07-02 report/paper.tex
@@
 \usepackage{multirow}
+\usepackage{multicol}
 \usepackage{subfig}
 \usepackage{alphalph}
+\usepackage{scrextend}
+\usepackage{xcolor}
 \renewcommand\thesubfigure{\alphalph{\value{subfigure}}}
@@
 \title{Pulsar detection - Machine learning and pattern recognition exam}
-\author{Alberto Baroso - s296520}
-\date{2022-07-13}
+\author{Alberto Baroso \\ s296520}
+\vspace{1cm}
+\date{2023-01-31}
 \maketitle
@@
     Thus pulsar search involves looking for periodic radio signals with large radio telescopes.
-    The goal of this project is to use machine learning techniques to solve the binary classification problem of detecting pulsars from the Pulsar dataset.
+    \\ \\
+    The goal of this project is to find the best machine learning techniques to solve the binary classification problem of detecting pulsars from the Pulsar dataset.
+    The analysis starts with the exploration of the dataset and its features, it proceeds with the application of dimensionality reduction and preprocessing techniques,
+    different kinds of models are then evaluated and their scores are calibrated, finally, experiments are carried out on the test dataset to evaluate if the choices made during training were optimal.
 \end{abstract}
@@
 
-\subsection{Features}
+\subsection{Datasets}
 
-Pulsar candidates are described by a class label and 8 continuous features extracted from radio signals collected by radio telescopes. \vspace{0.2cm}
+The Pulsar dataset \cite{stw656} contains 17.898 pulsar candidates of which 1.639 are real pulsars and 16,259 are spurious examples.
+The distribution of the two classes is highly unbalanced.
+The data has been split into a training set and a test set:
+\begin{itemize}
+    \item The training set contains 8929 samples, of which 8108 are non-pulsars and 821 are pulsars.
+    \item The test set contains 8969 samples, of which 8151 are non-pulsars and 818 are pulsars.
+\end{itemize}
+The class labels used are: 0 for non-pulsar samples and 1 for pulsars.
 
+\subsection{Feature descriptions}
+
+Pulsar candidates are described by 8 continuous features extracted from radio signals collected by radio telescopes.
 The first four are simple statistics obtained from the integrated pulse profile (folded profile).
@@
 The remaining four features are similarly obtained from the DM-SNR curve (again see \cite{Lyon2016} for more details).
-\vspace{0.2cm}
 The features are listed below:
@@
     \item Skewness of the DM-SNR curve.
-    \item Class
 \end{enumerate}
-
-The class labels used are 0 (non pulsar) and 1 (pulsar).
 
@@
 \begin{itemize}
-    \item Features have mainly irregular distributions
-    \item Few features, the ones for non pulsars in figures (a), (b) and (g), present distributions similar to Gaussians
-    \item Looking at the graph (b) it's possible to deduce that the "Standard Deviation of the integrated profile" could be one of the less informative features as the histograms of the two classes share one of the largest overlapping areas compared to the other graphs.
-But since it also has a good amount of non overlapping areas it can still provide useful information for classification purposes.
-    \item Grphs (c) and (d) show that "Excess of kurtosis of integrated profile" and "Skewness of integrated profile" could be among the best features for sample discrimination, as they have little overlapping areas of the histograms.
-    \item Some features, mainly the ones in graphs (b), (d) and (h), are characterized by the presence of significantly large outliers.
-Similar claims can be made by looking at the ranges of minimum and maximum values for the corresponding features
-    \item We can expect classification algorithms to provide less than ideal results due to the presence of outliers.
+    \item Features mainly follow irregular distributions
+    \item Few features, the ones for non-pulsars in figures (a), (b), and (g), present distributions similar to Gaussians
+    \item Looking at graph (b) it's possible to deduce that the "Standard Deviation of the integrated profile" could be one of the less informative features as the histograms of the two classes share one of the largest overlapping areas compared to the other graphs.
+But since it also has a good amount of non-overlapping areas it can still provide useful information for classification purposes.
+    \item Graphs (c) and (d) show that "Excess of kurtosis of integrated profile" and "Skewness of integrated profile" could be among the best features for sample discrimination, as they have little overlapping areas of the histograms.
+    \item Values for non-pulsars are generally concentrated around some values, while pulsars present a wider range of values.
+    \item Some features, mainly the ones in graphs (b), (d), and (h), are characterized by the presence of significantly large outliers.
+Similar claims can be made by looking at the ranges of minimum and maximum values for the corresponding features.
+It's possible to expect classification algorithms to provide less than ideal results due to the presence of outliers.
 \end{itemize}
 
-% As a result, we further pre-process the data by "Gaussianizing" the features.
-
-\subsection{Feature pairs}
+\subsection{Feature correlation}
 
@@
 
-% 
-% Classes are highly imbalanced.
-%
-
-Observing the feature pairs it's possible to note that there isn't one that allows to find a clear separation of the two classes. 
-\vspace*{0.5cm}
-The combination of features represented in graphs (n), (p), (ab), (af) and (ah) get close to identify regions of space where samples are well separated.
-\vspace*{0.5cm}\\
-The least useful pairs of features are represented in graphs (q), (s), (u) and (ac), where samples are almost entirely overlapping.
-
-% TODO: gaussianize features
-% TODO: feature correlation: compute pearson correlation heatmap
+Observing the feature pairs it's possible to note that there isn't one that allows finding a clear separation of the two classes. \\ \\
+The combination of features represented in graphs (n), (p), (ab), (af), and (ah) get close to identifying regions of space where samples are well separated. \\ \\
+The least useful pairs of features are represented in graphs (q), (s), (u), and (ac), where samples are almost entirely overlapping. \\ \\
+The plots also show that some pairs of features share similar distributions, implying that there is some sort of correlation between them. \\ \\
+Overall, it's reasonable to believe that there are enough pairs of features that separate classes well enough to obtain good accuracy for the trained models.
 
 \clearpage
+
+Pearson correlation \( \frac{Cov(X,Y)}{\sqrt{Var(X)}\sqrt{Var(Y)}} \) was used to explore the correlation between all pairs of features.
+The absolute value of the Pearson correlation between all pairs of features is shown in the following heatmaps: \\
+
+\begin{figure}[H]
+    \caption{Pearson correlation heatmap - RAW features}
+    \begin{center}
+        \begin{tabular}{ccc}
+            \hspace*{-65pt}
+            \subfloat[Whole dataset]{\includegraphics[width = 160pt]{img/pearson_correlation/raw_heatmap.png}}  &
+            \subfloat[Pulsars]{\includegraphics[width = 160pt]{img/pearson_correlation/raw_pulsar_heatmap.png}} &
+            \subfloat[Non-pulsars]{\includegraphics[width = 160pt]{img/pearson_correlation/raw_non_pulsar_heatmap.png}} \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\begin{figure}[H]
+    \caption{Pearson correlation heatmap - Z-normalized features}
+    \begin{center}
+        \begin{tabular}{ccc}
+            \hspace*{-65pt}
+            \subfloat[Whole dataset]{\includegraphics[width = 160pt]{img/pearson_correlation/z-normalization_heatmap.png}} &
+            \subfloat[Pulsars]{\includegraphics[width = 160pt]{img/pearson_correlation/z-normalization_pulsar.png}}        &
+            \subfloat[Non-pulsars]{\includegraphics[width = 160pt]{img/pearson_correlation/z-normalization_non_pulsar.png}}  \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+Both raw features and z-normalized features heatmaps highlight the presence of strong correlation between some pairs of features. In particular between features:
+
+\begin{enumerate}
+    \item (0) Mean of the integrated profile - (2) Excess kurtosis of the integrated profile.
+    \item (1) StDev of the integrated profile - (3) Skewness of the integrated profile.
+    \item (4) Mean of the DM-SNR curve - (5) Standard deviation of the DM-SNR curve.
+    \item (6) Excess kurtosis of the DM-SNR curve - (7) Skewness of the DM-SNR curve.
+\end{enumerate}
+
+In general, the first four features appear to be related to each other and this is because they are all measures of the integrated profile.
+The same consideration applies to the last four features, while the first four and the last four present little correlation.
+The presence of correlation between features suggests that some models, such as diagonal MVG classifiers, might benefit from the application of PCA to reduce the amount of features.
+Applying PCA effectively reduces the overall feature correlation, below the heatmap obtained after the application of PCA to 4 dimensions is shown: \\
+
+\begin{figure}[H]
+    \caption{Pearson correlation heatmap - PCA features}
+    \begin{center}
+        \begin{tabular}{ccc}
+            \hspace*{-65pt}
+            \subfloat[Whole dataset]{\includegraphics[width = 160pt]{img/pearson_correlation/pca_heatmap.png}} &
+            \subfloat[Pulsars]{\includegraphics[width = 160pt]{img/pearson_correlation/pca_pulsar.png}}        &
+            \subfloat[Non-pulsars]{\includegraphics[width = 160pt]{img/pearson_correlation/pca_non_pulsar.png}}  \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\clearpage
+
+\subsection{Applications}
+
+Given no prior knowledge of the environment in which the final model will work,
+3 different applications have been considered, one where the priors are balanced and two where the prior is biased towards one of the classes:
+\begin{itemize}
+    \item ( \(\tilde{\pi}\), Cfp, Cfn) = (0.5, 1, 1)
+    \item ( \(\tilde{\pi}\), Cfp, Cfn) = (0.1, 1, 1)
+    \item ( \(\tilde{\pi}\), Cfp, Cfn) = (0.9, 1, 1)
+\end{itemize}
+
+\section{Data preprocessing}
+
+\subsection{K-fold cross-validation}
+
+The K-fold approach has been used instead of the single-split validation technique to achieve more robust results, as the time required for K-fold is not excessive.
+Models are compared by means of the Minimum Detection Cost function, a measure of the cost paid for making optimal decisions on the validation set. \\
+
+The original training set has been split into 10 folds of similar size using the K-fold cross-validation technique.
+At each iteration of the K-fold cross-validation algorithm, 9 folds have been employed for training and 1 for validation.
+The data preprocessing techniques have been applied to the data after it was split into folds,
+i.e. the transformation to be applied, e.g. PCA projection matrix, has been computed on the 9 train folds, then it has been applied to both training and validation folds.
+This was done to avoid bias in the results, as splitting after preprocessing would have meant classifying data that the model had already seen.
+\\ \\The procedure followed to train each model is the same:
+\begin{enumerate}
+    \item Divide training samples into K-folds, eventually applying preprocessing techniques (Computation performed only once)
+    \item For each fold:
+          \begin{itemize}
+              \item Train a new model on the K-1 training folds
+              \item Use the trained model to score the validation fold
+          \end{itemize}
+    \item Combine the scores of all folds
+    \item Use the scores to compute minimum detection cost
+\end{enumerate}
+
+\subsection{Shuffling}
+
+Before applying any preprocessing technique, the training dataset has been randomly shuffled to avoid any bias that could be introduced by the order of the samples.
+
+\subsection{Z-Normalization}
+
+All samples have been processed by means of Z-normalization, i.e. each feature has been transformed to have zero mean and unit variance.
+$ x_i = \frac{x_i - \mu}{\sigma} $
+$\mu$ and $\sigma$ are respectively the mean and the standard deviation of each feature describing the samples.
+
+\subsection{Feature Gaussianization}
+
+As an alternative preprocessing step to z-normalization, raw features have been "Gaussianized", i.e. their distributions have been transformed to be more similar to Gaussian distributions.
+The Guasianization of a feature consists in computing a rank for each sample with respect to all samples and then computing the inverse of the cumulative distribution function of the standard normal distribution.
+The rank of both training and evaluation data is computed with respect to the training data only.
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \subfloat[]{\includegraphics[width = 200pt]{img/gaussianized_features/mean_of_ip}}                    &
+            \subfloat[]{\includegraphics[width = 200pt]{img/gaussianized_features/stdev_of_ip.png}}                 \\
+            \subfloat[]{\includegraphics[width = 200pt]{img/gaussianized_features/excess_kurtosis_of_ip.png}}     &
+            \subfloat[]{\includegraphics[width = 200pt]{img/gaussianized_features/skewness_of_ip.png}}              \\
+            \subfloat[]{\includegraphics[width = 200pt]{img/gaussianized_features/mean_of_dm-snr}}                &
+            \subfloat[]{\includegraphics[width = 200pt]{img/gaussianized_features/stdev_of_dm-snr.png}}             \\
+            \subfloat[]{\includegraphics[width = 200pt]{img/gaussianized_features/excess_kurtosis_of_dm-snr.png}} &
+            \subfloat[]{\includegraphics[width = 200pt]{img/gaussianized_features/skewness_of_dm-snr.png}}          \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+Gaussianization is computed over the whole training set, therefore class-conditional distributions might still not behave like a Gaussian distribution.
+This effect is clearly visible in the plots above, non-pulsars follow more regular distributions (due to their higher number of samples), while pulsars show in most cases moderate deviations from the Gaussian assumptions.
+Nonetheless, the new Gaussian-like distributions are likely to improve results of classifiers that assume data are Normal distributed.
+This preprocessing step also helps reducing the negative effects of the outliers as they have been mapped as "closer" values on the tail-end of the Gaussian distributions.
+
+\subsection{Linear Discriminant Analysis}
+Linear Discriminant Analysis (LDA) is deemed not useful as a preprocessing step for this dataset as LDA is capable of finding at most c - 1 discriminant directions, where c is the number of classes.
+Being this a binary classification problem, the maximum number of directions LDA can produce is 1.
+Given the huge loss of information this would imply, results obtained on LDA-preprocessed data will likely be worse than the ones of other approaches.
+Therefore, the use of LDA is not explored in this project.
+
+\subsection{Principal Component Analysis}
+
+The results from the correlation analysis suggest that some models might benefit from applying a dimensionality reduction technique such as PCA to obtain uncorrelated features.
+Removing correlation between features could also improve model performances by helping to reduce overfitting.
+Although, given the already small number of features, further reducing it might lead to a loss of information higher than the gain obtained by the removal of correlation.
+The PCA projection matrix has been computed on the training dataset and applied to the training and evaluation datasets.
+
+\clearpage
+
+\begin{multicols}{2}
+    {
+        \noindent
+        \begin{addmargin}[0em]{1em}
+            The effects of different values of the hyperparameter m (number of components) for PCA have been evaluated by means of minimum DCF over different MVG models.
+            Table 2 shows the results obtained with different values of m, types of Gaussian classifiers, and applications.
+            Full diagonal models performances worsen as the number of components decreases but overall this deterioration is not excessive.
+            Classifiers with diagonal covariance on the other hand slightly improve their performances as correlated features are removed, but their overall performance is worse than the full covariance models, meaning that the training dataset is large enough to reliably estimate the full covariance matrix and effectively capture the correlation in the data.
+            Since full covariance models are capable of capturing these correlations it doesn't bring any advantage to reduce the dimensionality of the data.
+PCA is therefore not used in the following experiments.
+        \end{addmargin}
+    }
+    {
+        \noindent
+        \renewcommand{\arraystretch}{1.2}
+        \begin{tabular}{cccc}
+
+            Model      & \(\tilde{\pi}\)=0.5  & \(\tilde{\pi}\)=0.1   & \(\tilde{\pi}\)=0.9   \\ [0.5ex]
+
+            \hline
+            \multicolumn{4}{c}{RAW Features - No PCA}                                         \\
+            \hline
+            Full Cov.  & 0.141                & 0.280                 & 0.680                 \\
+            Diagonal   & 0.194                & 0.314                 & 0.723                 \\
+            Tied Full  & {\color{red} 0.113 } & {\color{blue} 0.224 } & {\color{blue} 0.571 } \\
+            Tied Diag. & 0.161                & 0.266                 & 0.591                 \\
+
+            \hline
+            \multicolumn{4}{c}{RAW Features - PCA (m = 7)}                                    \\
+            \hline
+            Full Cov.  & 0.174                & 0.302                 & 0.708                 \\
+            Diagonal   & 0.220                & 0.628                 & 0.833                 \\
+            Tied Full  & {\color{red} 0.153 } & {\color{blue} 0.280 } & {\color{blue} 0.612 } \\
+            Tied Diag. & 0.172                & 0.306                 & 0.639                 \\
+
+            \hline
+            \multicolumn{4}{c}{RAW Features - PCA (m = 6)}                                    \\
+            \hline
+            Full Cov.  & 0.172                & 0.315                 & 0.662                 \\
+            Diagonal   & 0.220                & 0.645                 & 0.818                 \\
+            Tied Full  & {\color{red} 0.156 } & {\color{blue} 0.288 } & {\color{blue} 0.602 } \\
+            Tied Diag. & 0.173                & 0.310                 & 0.618                 \\
+
+            \hline
+            \multicolumn{4}{c}{RAW Features - PCA (m = 5)}                                    \\
+            \hline
+            Full Cov.  & 0.183                & 0.423                 & 0.748                 \\
+            Diagonal   & 0.203                & 0.607                 & 0.797                 \\
+            Tied Full  & {\color{red} 0.179 } & {\color{blue} 0.315 } & {\color{blue} 0.592 } \\
+            Tied Diag. & 0.187                & 0.345                 & 0.604                 \\
+
+            \hline
+            \multicolumn{4}{c}{RAW Features - PCA (m = 4)}                                    \\
+            \hline
+            Full Cov.  & 0.186                & 0.434                 & 0.683                 \\
+            Diagonal   & 0.199                & 0.603                 & 0.799                 \\
+            Tied Full  & {\color{red} 0.172 } & {\color{blue} 0.316 } & {\color{blue} 0.601 } \\
+            Tied Diag. & 0.184                & 0.347                 & 0.611                 \\
+        \end{tabular}
+        \captionof{table}{MVG Classifiers - Tuning of PCA components}
+    }
+\end{multicols}
+
+\section{Classification}
+
+\subsection{Gaussian models}
+
+\subsubsection{Multivariate Gaussian Classifier}
+The MVG classifier uses full covariance matrices, i.e. a covariance matrix and a mean is estimated for each classes.
+Given the large number of training samples, this kind of model is expected to provide good results compared to the other types of Gaussian classifiers.
+
+\subsubsection{Tied Gaussian Classifier}
+The tied MVG model uses the same covariance matrix for all classes.
+This assumption shouldn't benefit the pulsar classification problem as the training set contains a good number of samples allowing covariance matrices to be reliably estimated for each class.
+
+\subsubsection{Naive Bayes Gaussian Classifier}
+The naive Bayes MVG model simplifies the estimation by assuming that the features are independent.
+We thus expect the MVG classifier with naive Bayes assumption to perform worse that the previously seen
+models because the correlation heatmaps highlighted a strong correlation between features.
+
+\subsubsection{Tied Naive Bayes Gaussian Classifier}
+The tied naive Bayes MVG combines the assumptions of feature independence and tied covariance matrices.
+Given that the Naive assumption is not met, this model is also expected to have some among the worst performances.
+
+\clearpage
+
+\subsubsection{Results}
+
+\begin{multicols}{2}
+    {
+
+        \begin{addmargin}[0em]{1em}
+            As expected Diagonal models obtain the worst performances.
+            Contrary to expectations though, tied covariance models provide better results than full covariance ones.
+            This could be due to the fact that since the classes represent the same physical phenomena the covariance matrices of the two classes might be similar.
+            Another possible explanation could be that linear separation rules, such as the one found by the tied model, are more appropriate than quadratic ones for this classification problem.
+
+            Raw and Z-normalized features perform similarly.
+            Gaussianized features improve the performances of the diagonal covariance model but overall perform worse than the Raw and Z-normalized features on the other models.
+
+            \(\tilde{\pi}\) = 0.9 produces the worst results, \(\tilde{\pi}\) = 0.1 provides mediocre outputs but the best ones are obtained for application \(\tilde{\pi}\) = 0.5.
+            The best performances are given by the Tied Full Covariance model on the Z-Normalized Features.
+        \end{addmargin}
+    }
+    {
+        \noindent
+        \renewcommand{\arraystretch}{1.2}
+        \begin{tabular}{@{}cccc@{}}
+            Model      & \(\tilde{\pi}\)=0.5  & \(\tilde{\pi}\)=0.1   & \(\tilde{\pi}\)=0.9  \\ [0.5ex]
+
+            \hline
+            \multicolumn{4}{c}{RAW Features}                                                 \\
+            \hline
+            Full Cov.  & 0.141                & 0.280                 & 0.680                \\
+            Diagonal   & 0.194                & 0.314                 & 0.723                \\
+            Tied Full  & {\color{red} 0.113 } & {\color{red} 0.224 }  & 0.571                \\
+            Tied Diag. & 0.161                & 0.267                 & 0.591                \\
+
+            \hline
+            \multicolumn{4}{c}{Z-Normalized Features}                                        \\
+            \hline
+            Full Cov.  & 0.141                & 0.280                 & 0.680                \\
+            Diagonal   & 0.194                & 0.314                 & 0.723                \\
+            Tied Full  & {\color{red} 0.113 } & {\color{red} 0.224 }  & 0.571                \\
+            Tied Diag. & 0.161                & 0.267                 & 0.591                \\
+
+            \hline
+            \multicolumn{4}{c}{Gaussianized Features}                                        \\
+            \hline
+            Full Cov.  & 0.154                & 0.245                 & 0.717                \\
+            Diagonal   & 0.152                & 0.277                 & 0.602                \\
+            Tied Full  & 0.132                & {\color{blue} 0.237 } & {\color{red} 0.532 } \\
+            Tied Diag. & 0.163                & 0.292                 & 0.622                \\
+        \end{tabular}
+        \captionof{table}{MVG Classifiers - Comparison of results}
+    }
+\end{multicols}
+
+\subsection{Logistic Regression}
+
+The logistic regression model is trained by minimizing the sum of the logistic loss over all samples.
+The L-BFGS algorithm has been used as numerical solver to find the minimum.
+The logistic regression model makes no assumptions on data distribution but requires a hypothesis
+on the shape of the separation rule to find. A linear separation surface has been assumed: \[ w^T x+b \]
+Where w is a vector perpendicular to the separation surface and b is a bias that tells where on w the separation rule lies.
+
+To handle class unbalance, a prior-weighted regularized version of the objective function has been used: \\
+
+\begingroup{
+\setlength{\thinmuskip}{0mu}
+\hspace*{-30pt}
+$J(w, b)=\frac{\lambda}{2}\|w\|^2+\frac{\pi_T}{n_T} \sum\limits_{i=\left.1\right| c_i=1 }^n \log \left(1+e^{-z_i (w^Tx_i+b)}\right)+\frac{1-\pi_T}{n_F} \sum\limits_{i=\left.1\right| c_i=0 }^n \log \left(1+e^{-z_i (w^Tx_i+b)}\right)$ \\ \\
+}\endgroup
+\(\lambda\) is the regularization coefficient, \(\pi_T\) is the prior probability of the positive class, \(n_T\) is the number of positive samples, \(n_F\) is the number of negative samples, \(c_i\) is the class of the \(i\)-th sample.
+
+\subsubsection{Hyperparameter tuning: \(\lambda\)}
+
+Small values of \(\lambda\) lead to poor generalization, while higher values of \(\lambda\) allow finding small values of \(\|w\|\) but worse separation rules.
+To find the best value of \(\lambda\) the minDCF has been plotted for different applications and values of \(\lambda\) on a logarithmic scale.
+For the purpose of evaluating different values of \(\lambda\), only the main application was considered (\(\pi_T = \frac{1}{2}\))
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/evaluation_plots/lambda-lr-z-normalized.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/lambda-lr-gaussianized.png}   \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+The best results are obtained with \(\lambda = 0 \) for all the applications, and both Z-Normalized Gaussianized features.
+The reason why regularization is not helpful can be found in the small number of dimensions, which implies little risk of overfitting.
+Even if regularization provides little to no improvements, in the following experiments \(\lambda\) has been set to \(10^{-6}\).
+
+\clearpage
+
+\subsubsection{Results}
+
+\begin{center}
+    \renewcommand{\arraystretch}{1.2}
+    \begin{tabular}{@{}cccc@{}}
+        Model                                                    & \(\tilde{\pi}\)=0.5   & \(\tilde{\pi}\)=0.1  & \(\tilde{\pi}\)=0.9   \\ [0.5ex]
+
+        \hline
+        \multicolumn{4}{c}{RAW Features}                                                                                                \\
+        \hline
+        Logisitc Regression \((\lambda = 10^{-6}, \pi_T = 0.5)\) & {\color{blue} 0.116 } & 0.219                & {\color{blue} 0.530 } \\
+        Logisitc Regression \((\lambda = 10^{-6}, \pi_T = 0.1)\) & {\color{blue} 0.114 } & {\color{red} 0.213 } & 0.538                 \\
+        Logisitc Regression \((\lambda = 10^{-6}, \pi_T = 0.9)\) & {\color{blue} 0.118 } & 0.218                & 0.534                 \\
+
+        \hline
+        \multicolumn{4}{c}{Z-normalized Features}                                                                                       \\
+        \hline
+        Logisitc Regression \((\lambda = 10^{-6}, \pi_T = 0.5)\) & {\color{blue} 0.116 } & 0.219                & {\color{blue} 0.530 } \\
+        Logisitc Regression \((\lambda = 10^{-6}, \pi_T = 0.1)\) & {\color{red} 0.113 }  & {\color{red} 0.213 } & 0.537                 \\
+        Logisitc Regression \((\lambda = 10^{-6}, \pi_T = 0.9)\) & 0.120                 & 0.220                & 0.537                 \\
+
+        \hline
+        \multicolumn{4}{c}{Gaussianized Features}                                                                                       \\
+        \hline
+        Logisitc Regression \((\lambda = 10^{-6}, \pi_T = 0.5)\) & 0.126                 & 0.238                & 0.534                 \\
+        Logisitc Regression \((\lambda = 10^{-6}, \pi_T = 0.1)\) & 0.128                 & 0.229                & 0.537                 \\
+        Logisitc Regression \((\lambda = 10^{-6}, \pi_T = 0.9)\) & 0.127                 & 0.238                & {\color{red} 0.524 }  \\
+
+        \hline
+        \multicolumn{4}{c}{Other models}                                                                                                \\
+        \hline
+        MVG Tied Full Covariance                                 & {\color{red} 0.113 }  & 0.224                & 0.571                 \\
+    \end{tabular}
+    \captionof{table}{Logistic Regression - Comparison of results}
+\end{center}
+
+Gaussianized features perform worse than Z-normalized and Raw features, as Logistic Regression doesn't require assumptions on the distribution of the data.
+Table 4 shows that changing the value of \(\pi_T\) has little effect on the results.
+Once again \(\tilde{\pi}\) = 0.9 performs worse than \(\tilde{\pi}\) = 0.1, which in turn has worse performances than \(\tilde{\pi}\) = 0.5.
+Best results are obtained for Z-normalized features, with \(\pi_T = 0.1\) and \(\tilde{\pi} = 0.5\).
+The improvement of using \(\pi_T = 0.1\) instead of \(\pi_T = 0.5\) is very small, but this can be explained by the unbalance in the dataset towards non-pulsar samples.
+Tied MVG with full covariance has similar performances to the best logistic regression model.
+
+\subsection{Support Vector Machines}
+
+The SVM problem was solved using its dual formulation: \\
+
+\begin{center}
+    $ \max_{\alpha} \alpha^T1-\frac{1}{2}\alpha^TH\alpha $ subject to $ 0\leq\alpha_i\leq C_i, i=1,..,n $
+\end{center}
+
+An updated version of the constraints has also been considered to handle the unbalance between the classes.
+Constraints have been changed to use costs proportional to the priors defined by the application:
+$ C_T = C \cdot \frac{\pi_T}{\pi_T^{emp}} $
+$ C_F = C \cdot \frac{\pi_F}{\pi_F^{emp}} $
+where $ C_i = C_T $ for pulsars and $ C_i = C_F $ for non-pulsar samples.
+$\pi_T^{emp}$ and $\pi_F^{emp}$ are the empirical priors obtained from the training set.
+
+\subsubsection{Hyperparameter tuning: C}
+C represents the tradeoff between having good accuracy on the training set and finding a large margin (better generalization).
+To find the best value of C, K-fold cross-validation has been employed to plot the minDCF for different applications and different values of C on a logarithmic scale:
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/evaluation_plots/C-svm-z-normalized-unbalanced.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/C-svm-z-normalized-balanced.png}     \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/evaluation_plots/C-svm-gaussianized-unbalanced.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/C-svm-gaussianized-balanced.png}     \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+Low values of C show the models start to underfit, while high values of C clearly show signs of overfitting.
+The best results are obtained with C $\approx 10^{-1}$ for both balanced ($\pi_T = 0.5$) and unbalanced applications, and for both z-normalized and gaussianized features.
+
+\subsubsection{Results}
+
+Minimum DCF has been computed for different values of \(\pi_T\) and \(\tilde{\pi}\) for both z-normalized and gaussianized features.
+Where $ \pi_T $ is the prior probability for the true hypothesis with which the model is trained and $ \tilde{\pi} $ represents the effective prior in the context where the model will be used.
+
+\begin{center}
+    \renewcommand{\arraystretch}{1.2}
+    \begin{tabular}{@{}cccc@{}}
+        Model                                        & \(\tilde{\pi}\)=0.5   & \(\tilde{\pi}\)=0.1  & \(\tilde{\pi}\)=0.9  \\ [0.5ex]
+
+        \hline
+        \multicolumn{4}{c}{RAW Features}                                                                                   \\
+        \hline
+        Linear SVM (C = $10^{-1}$, unbalanced)       & 0.629                 & 0.803                & 0.954                \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.5$)    & 0.592                 & 0.972                & 0.936                \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.1$)    & 0.714                 & 1.000                & 0.931                \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.9$)    & 0.599                 & 0.847                & 0.939                \\
+
+        \hline
+        \multicolumn{4}{c}{Z-normalized Features}                                                                          \\
+        \hline
+        Linear SVM (C = $10^{-1}$, unbalanced)       & 0.117                 & 0.222                & 0.584                \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.5$)    & {\color{blue} 0.113 } & 0.223                & 0.558                \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.1$)    & {\color{red} 0.111 }  & {\color{red} 0.214 } & 0.570                \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.9$)    & 0.128                 & 0.234                & 0.527                \\
+
+        \hline
+        \multicolumn{4}{c}{Gaussianized Features}                                                                          \\
+        \hline
+        Linear SVM (C = $10^{-1}$, unbalanced)       & 0.135                 & 0.239                & 0.531                \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.5$)    & 0.120                 & 0.251                & {\color{red} 0.519 } \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.1$)    & 0.133                 & 0.237                & 0.531                \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.9$)    & 0.130                 & 0.244                & 0.525                \\
+
+        \hline
+        \multicolumn{4}{c}{Other models}                                                                                   \\
+        \hline
+        MVG Tied Full Cov.                           & 0.113                 & 0.224                & 0.571                \\
+        Log Reg \((\lambda = 10^{-6}, \pi_T = 0.1)\) & 0.113                 & 0.213                & 0.537                \\
+    \end{tabular}
+    \captionof{table}{Linear SVM - Comparison of results}
+\end{center}
+
+Table 5 shows that balancing classes has little impact on the outcome.
+Raw features are the worst-performing ones, while z-normalized features perform better than gaussianized ones.
+Best results are obtained for the z-normalized features with $\pi_T = 0.1$, \(\tilde{\pi}\) = 0.5, these are very close to the ones obtained for the best MVG and Logistic Regression models.
+
+\subsubsection{Support Vector Machines with Polynomial Kernel}
+
+The polynomial kernel $K(x_i, x_j) = (x_i^Tx_j + c)^d$ was used to train a linear separation surface in the expanded feature space, which corresponds to a quadratic separation rule in the original feature space.
+Hyperparameters C and c were jointly optimized using a grid search, d was set to 2.
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/evaluation_plots/polynomial-svm-z-normalized-C-c.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/polynomial-svm-gaussianized-C-c.png}   \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+\vspace{-20pt}
+\noindent
+The polynomial kernel yields the best results when used with C=1 and c=0.1 for the z-normalized features and C=1, c=1 for the gaussianized features. \\
+
+\renewcommand{\arraystretch}{1.2}
+\begin{tabular}{@{}cccc@{}}
+    Model                                      & \(\tilde{\pi}\)=0.5  & \(\tilde{\pi}\)=0.1  & \(\tilde{\pi}\)=0.9  \\ [0.5ex]
+
+    \hline
+    \multicolumn{4}{c}{Z-normalized Features}                                                                       \\
+    \hline
+    Polynomial Kernel (c = 0.1, Unbalanced)    & 0.133                & 0.246                & 0.666                \\
+    Polynomial Kernel (c = 0.1, $\pi_T = 0.5$) & {\color{red} 0.103 } & {\color{red} 0.220 } & {\color{red} 0.525 } \\
+    Polynomial Kernel (c = 0.1, $\pi_T = 0.1$) & 0.129                & 0.228                & 0.662                \\
+    Polynomial Kernel (c = 0.1, $\pi_T = 0.9$) & 0.141                & 0.297                & 0.594                \\
+
+    \hline
+    \multicolumn{4}{c}{Gaussianized Features}                                                                       \\
+    \hline
+    Polynomial Kernel (c = 1, Unbalanced)      & 0.147                & 0.262                & 0.638                \\
+    Polynomial Kernel (c = 1, $\pi_T = 0.5$)   & 0.128                & 0.255                & 0.570                \\
+    Polynomial Kernel (c = 1, $\pi_T = 0.1$)   & 0.136                & 0.244                & 0.634                \\
+    Polynomial Kernel (c = 1, $\pi_T = 0.9$)   & 0.207                & 0.485                & 0.603                \\
+\end{tabular}
+\captionof{table}{Polynomial Kernel SVM - Comparison of results}
+
+\vspace*{10pt}
+The best results are obtained for the Z-normalized features, with c = 0.1, \(\pi_T = 0.5\), and \(\tilde{\pi}\)=0.5.
+The polynomial Kernel finds better results than any previous model but the difference is not significant.
+
+\subsubsection{Support Vector Machines with RBF Kernel}
+
+The RBF kernel $K(x_i, x_j) = e^{-\gamma||x_i - x_j||^2}$ requires estimating the hyperparameter $\gamma$ jointly with the hyperparameter C, this was performed using a grid search.
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/evaluation_plots/rbf-svm-z-normalized-C-gamma.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/rbf-svm-gaussianized-C-gamma.png}   \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\vspace{-20pt}
+\noindent
+The RBF kernel has been used with C=$10^2$, $\gamma=0.1$ for both z-normalized and Gaussianized features. \\
+
+\begin{center}
+    \renewcommand{\arraystretch}{1.2}
+    \begin{tabular}{@{}cccc@{}}
+        Model                                    & \(\tilde{\pi}\)=0.5   & \(\tilde{\pi}\)=0.1   & \(\tilde{\pi}\)=0.9  \\ [0.5ex]
+
+        \hline
+        \multicolumn{4}{c}{Z-normalized Features}                                                                       \\
+        \hline
+        RBF Kernel ($\gamma$ = 0.1, Unbalanced)  & 0.121                 & {\color{blue} 0.211 } & 0.867                \\
+        RBF Kernel ($\gamma = 0.1, \pi_T = 0.5$) & {\color{blue} 0.115 } & 0.236                 & 0.642                \\
+        RBF Kernel ($\gamma = 0.1, \pi_T = 0.1$) & 0.123                 & {\color{blue} 0.211 } & 0.830                \\
+        RBF Kernel ($\gamma = 0.1, \pi_T = 0.9$) & 0.220                 & 0.690                 & 0.630                \\
+
+        \hline
+        \multicolumn{4}{c}{Gaussianized Features}                                                                       \\
+        \hline
+        RBF Kernel ($\gamma$ = 0.1, Unbalanced)  & 0.120                 & {\color{red} 0.206 }  & 0.770                \\
+        RBF Kernel ($\gamma = 0.1, \pi_T = 0.5$) & {\color{red} 0.109 }  & 0.219                 & 0.631                \\
+        RBF Kernel ($\gamma = 0.1, \pi_T = 0.1$) & 0.121                 & {\color{red} 0.206 }  & 0.773                \\
+        RBF Kernel ($\gamma = 0.1, \pi_T = 0.9$) & 0.201                 & 0.460                 & {\color{red} 0.604 } \\
+    \end{tabular}
+    \captionof{table}{RBF Kernel SVM - Comparison of results}
+\end{center}
+
+The best results are obtained for the Gaussianized features, with $\gamma$ = 0.1, \(\pi_T = 0.5\), and \(\tilde{\pi}\)=0.5.
+Using different priors is effective in providing somewhat better results for different applications.
+The RBF Kernel finds slightly worse results than the polynomial kernel.
+
+\subsection{Gaussian Mixture Models}
+
+GMM is a generative approach that can approximate generic distributions.
+The GMMs have been estimated by iteratively applying the LBG algorithm to duplicate the number of components and the Expectation Maximization algorithm to estimate the parameters of each component.
+
+\subsubsection{Hyperparameter tuning: number of components}
+
+Full-covariance, diagonal-covariance, and tied-covariance models have been considered.
+The performances of models with a varying number of components have been evaluated using the minDCF metric through K-fold cross-validation.
+For simplicity only application $\tilde{\pi}$=0.5 has been considered.
+
+\begin{figure}[H]
+    \caption{GMM Components Evaluation}
+    \begin{center}
+        \begin{tabular}{ccc}
+            \hspace*{-65pt}
+            \includegraphics[width = 160pt]{img/evaluation_plots/gmm-components-full-covariance.png} &
+            \includegraphics[width = 160pt]{img/evaluation_plots/gmm-components-tied-covariance.png} &
+            \includegraphics[width = 160pt]{img/evaluation_plots/gmm-components-diagonal-covariance.png} \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+The model with full covariance matrix shows signs of overfitting for 16 components and above, while tied and diagonal covariance models are capable of handling a higher number of components.
+Experiments show that the best number of components is: 8 for the full-covariance model and 32 for both the tied-covariance and diagonal-covariance models.
+Z-normalized features obtain better results than gaussianized ones.
+
+\subsubsection{Results}
+
+Minimum DCF values have been computed for: different applications, feature preprocessing, and different types of models, using the best number of components found in the previous section.
+
+\begin{center}
+    \renewcommand{\arraystretch}{1.2}
+    \begin{tabular}{@{}cccc@{}}
+        Model                                        & \(\tilde{\pi}\)=0.5  & \(\tilde{\pi}\)=0.1   & \(\tilde{\pi}\)=0.9  \\ [0.5ex]
+
+        \hline
+        \multicolumn{4}{c}{RAW Features}                                                                                   \\
+        \hline
+        GMM Full Cov. 8 components                   & 0.130                & 0.245                 & 0.630                \\
+        GMM Tied Cov. 32 components                  & 0.147                & 0.297                 & 0.715                \\
+        GMM Diag. Cov. 32 components                 & 0.162                & 0.288                 & 0.659                \\
+        GMM Tied Diag. Cov. 32 components            & 0.159                & 0.300                 & 0.691                \\
+
+        \hline
+        \multicolumn{4}{c}{Z-normalized Features}                                                                          \\
+        \hline
+        GMM Full Cov. 8 components                   & {\color{red} 0.118 } & {\color{red} 0.224 }  & {\color{red} 0.562 } \\
+        GMM Tied Cov. 32 components                  & 0.129                & 0.250                 & 0.615                \\
+        GMM Diag. Cov. 32 components                 & 0.142                & 0.276                 & 0.599                \\
+        GMM Tied Diag. Cov. 32 components            & 0.153                & 0.291                 & 0.647                \\
+
+        \hline
+        \multicolumn{4}{c}{Gaussianized Features}                                                                          \\
+        \hline
+        GMM Full Cov. 8 components                   & 0.128                & 0.230                 & 0.680                \\
+        GMM Tied Cov. 32 components                  & 0.139                & {\color{blue} 0.226 } & 0.668                \\
+        GMM Diag. Cov. 32 components                 & 0.151                & 0.284                 & 0.662                \\
+        GMM Tied Diag. Cov. 32 components            & 0.184                & 0.293                 & 0.672                \\
+
+        \hline
+        \multicolumn{4}{c}{Other models}                                                                                   \\
+        \hline
+        MVG Tied Full Cov.                           & 0.113                & 0.224                 & 0.571                \\
+        Log Reg \((\lambda = 10^{-6}, \pi_T = 0.1)\) & 0.113                & 0.213                 & 0.537                \\
+        Linear SVM (C = $10^{-1}$, $\pi_T = 0.1$)    & 0.111                & 0.214                 & 0.570                \\
+    \end{tabular}
+    \captionof{table}{GMM - Comparison of results}
+\end{center}
+
+Diagonal models perform worse than any other model, due to the dataset being large enough to reliably estimate full covariance matrices.
+Even if using more components, tied models perform worse than non-tied full covariance models, this could be due to data being split into a few large clusters rather than many small ones.
+As confirmed by all other previous approaches, \(\tilde{\pi}\) = 0.5 yields better results than \(\tilde{\pi}\) = 0.1 and \(\tilde{\pi}\) = 0.9.
+Overall z-normalized features produce the best results, this could be due to the reduced discriminability of clusters after applying Gaussianization.
+Best results are obtained for the full covariance model with 8 components, z-normalized features, and $\tilde{\pi} = 0.5$.
+The best-performing GMM model finds slightly worse results than any other model previously evaluated.
+
+\subsection{Model selection}
+
+The experiments highlighted that the best performing models are:
+{\noindent
+\begin{tabular}{@{}cccc@{}}
+    Model                                                  & \(\tilde{\pi}\)=0.5 & \(\tilde{\pi}\)=0.1 & \(\tilde{\pi}\)=0.9 \\
+    \hline
+    MVG Tied Full Covariance                               & 0.113               & 0.224               & 0.571               \\
+    Linear Log Reg $(\lambda = 10^{-6}, \pi_T = 0.1)$      & 0.113               & 0.213               & 0.537               \\
+    Linear SVM (C = $10^{-1}$, $\pi_T = 0.1$)              & 0.111               & 0.214               & 0.570               \\
+    Polynomial kernel SVM (C = 1, c = 0.1, $\pi_T = 0.5$)  & 0.103               & 0.220               & 0.525               \\
+    RBF kernel SVM (C=$10^2$, $\gamma=0.1$, $\pi_T = 0.5$) & 0.109               & 0.219               & 0.631               \\
+    GMM Full Cov. 8 components                             & 0.118               & 0.224               & 0.562               \\
+\end{tabular}
+}
+
+If a single model had to be chosen based on these results, SVM with polynomial kernel would be the best choice, as it yields the best results across various applications.
+However, given the additional complexity required by the quadratic model and the minimal difference in performance, Linear Logistic regression is to be preferred due to its lower complexity.
+The trade-off between False Negative Rate and False Positive Rate has been plotted in the DET graph below:
+
+\begin{center}
+    \includegraphics[width=300pt]{img/evaluation_plots/DET_plot_best_models.png}
+\end{center}
+
+The Linear SVM performs better than the other models for low values of FPR, while the SVM with polynomial kernel is to be preferred if a small FNR is required.
+
+\clearpage
+
+Models have been so far evaluated using minimum DCF, which measures the cost of performing optimal decisions.
+Now the actual cost of the decisions taken by the classifiers is measured using the Actual DCF metric.
+To evaluate the difference between the two metrics, Bayes error plots have been computed for the selected models:
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/bayes/bayes-error-plot-mvg.png} &
+            \includegraphics[width = 200pt]{img/bayes/bayes-error-plot-lr.png}    \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/bayes/bayes-error-plot-polynomial-kernel-svm.png} &
+            \includegraphics[width = 200pt]{img/bayes/bayes-error-plot-rbf-kernel-svm.png}         \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/bayes/bayes-error-plot-svm.png} &
+            \includegraphics[width = 200pt]{img/bayes/bayes-error-plot-gmm.png}   \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+The Gaussian Mixture Model is already well-calibrated for a large number of applications.
+The MVG, SVM with polynomial kernel, and Logistic regression models present small but significant miscalibration for negative values of prior log-odds and large miscalibration errors for positive values of prior log-odds.
+The linear SVM and the SVM with RBF kernel are largely miscalibrated over the whole range of applications with the only exception of small positive values of the prior log odds for the SVM models using a kernel.
+
+\subsection{Score recalibration}
+
+Scores of each model have been recalibrated by training a model with probabilistic interpretation using the scores as a feature.
+A prior-weighted version of Logistic Regression has been used to perform the recalibration, $\tilde{\pi} = 0.5$ was used as the target application as this is expected to provide good calibration results for a wide range of applications.
+Given the low risk of overfitting due to the presence of a single dimension (the scores themselves), regularization has not been employed, i.e. $\lambda = 0$.
+The training set was split using the K-fold approach, at each iteration a LR model has been trained and used to score the validation fold, the scores have been then recalibrated by subtracting the prior log-odds of the target application, $\log{(\frac{\tilde{\pi}}{1-\tilde{\pi}})}$.
+Below the results obtained using calibrated scores are shown.
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/bayes/recalibrated-mvg.png} &
+            \includegraphics[width = 200pt]{img/bayes/recalibrated-lr.png}    \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/bayes/recalibrated-polynomial-kernel-svm.png} &
+            \includegraphics[width = 200pt]{img/bayes/recalibrated-rbf-kernel-svm.png}          \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/bayes/recalibrated-svm.png} &
+            \includegraphics[width = 200pt]{img/bayes/recalibrated-gmm.png}   \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+Overall the recalibration process has improved the performace of the models for a wide range of applications, only high values of prior log-odds are still slightly miscalibrated.
+
+\section{Experimental results}
+
+Data has been pre-processed with Z-normalization and Gaussianization techniques.
+The two transformations have been computed on the training set and then applied to both training and test sets.
+
+The best-performing models on the validation data have been re-trained using the entire training set to evaluate their behavior on the test set.
+For each model the following steps have been performed:
+\begin{itemize}
+    \item The training set was split with K-fold cross-validation
+    \item Each validation fold has been scored using a model trained on the training folds (preprocessing techniques have been applied at each iteration)
+    \item The whole set of produced scores has been used as a feature to train a prior-weighted Logistic Regression model for the application $\tilde{\pi} = 0.5$
+    \item The LR model was used to estimate a linear transformation of the scores which corresponds to the log-likelihood ratio of the two classes:
+          $ f(s) = \alpha s + \beta = \log{\frac{f_{S|C}(s|H_1)}{f_{S|C}(s|H_0)}}$
+    \item The original classifier has then been trained on the whole training set
+    \item The test samples have been scored by the classifier
+    \item The score of each test sample has been recalibrated as:\\ $s_{calibrated} = \alpha s + \beta = \alpha s + \beta' - \log{(\frac{\tilde{\pi}}{1-\tilde{\pi}})}$
+\end{itemize}
+
+Minimum DCFs and Actual DCFs have been computed for each model to evaluate if the proposed models are the best ones also for the test set. \\
+
+{
+\hspace{-58pt}
+\begin{tabular}{@{}c@{}c@{\hspace{3pt}}c@{}@{}c@{\hspace{3pt}}c@{}@{}c@{\hspace{3pt}}c@{}}
+    Model                                         & MinDCF                                  & ActDCF                                  & MinDCF                                  & ActDCF & MinDCF & ActDCF \\
+    \hline
+                                                  & \multicolumn{2}{c}{$\tilde{\pi} = 0.5$} & \multicolumn{2}{c}{$\tilde{\pi} = 0.1$} & \multicolumn{2}{c}{$\tilde{\pi} = 0.9$}                            \\
+    \hline
+    MVG Tied Full Covariance                      & 0.110                                   & 0.115                                   & 0.207                                   & 0.222  & 0.591  & 0.613  \\
+    Linear Log Reg $(\lambda=10^{-6}, \pi_T=0.1)$ & 0.110                                   & 0.113                                   & 0.199                                   & 0.219  & 0.530  & 0.574  \\
+    Linear SVM (C=$10^{-1}$, $\pi_T=0.1$)         & 0.112                                   & 0.116                                   & 0.202                                   & 0.215  & 0.549  & 0.576  \\
+    Polynomial SVM (C=1, c=0.1, $\pi_T=0.5$)      & 0.105                                   & 0.108                                   & 0.199                                   & 0.215  & 0.430  & 0.461  \\
+    RBF SVM (C=$10^2$, $\gamma=0.1$, $\pi_T=0.5$) & 0.111                                   & 0.112                                   & 0.231                                   & 0.246  & 0.574  & 0.658  \\
+    GMM Full Cov.
+8 components                    & 0.110                                   & 0.114                                   & 0.218                                   & 0.223  & 0.586  & 0.601  \\
+\end{tabular}
+\captionof{table}{Results on the test set}
+}
+
+\vspace*{20pt}
+All models achieve very good performances for applications $\tilde{\pi} = 0.5$ and $\tilde{\pi} = 0.1$.
+Overall the results are consistent with the observations made on the validation set.
+SVM with polynomial kernel is confirmed as the best model,
+while it provides minimal improvements with respect to other models for the balanced application,
+its superiority is more evident in unbalanced environments such as $\tilde{\pi} = 0.9$.
+
+\subsection{Multivariate Gaussian classifiers}
+
+To evaluate if the proposed solution is the best one also for the test set,
+Minimum and Actual DCFs have been computed on the test and compared with the results obtained for the validation set.
+
+\begin{center}
+    \renewcommand{\arraystretch}{1.2}
+    \begin{tabular}{c|ccc|ccc}
+
+        Model      & \(\tilde{\pi}\)=0.5            & \(\tilde{\pi}\)=0.1      & \(\tilde{\pi}\)=0.9  & \(\tilde{\pi}\)=0.5  & \(\tilde{\pi}\)=0.1  & \(\tilde{\pi}\)=0.9   \\ [0.5ex]
+
+        \hline
+                   & \multicolumn{3}{c|}{Validation} & \multicolumn{3}{c}{Test}                                                                                              \\
+
+        \hline
+        \multicolumn{7}{c}{RAW Features - No PCA}                                                                                                                           \\
+        \hline
+        Full Cov.  & 0.141                          & 0.280                    & 0.680                & 0.140                & 0.283                & 0.646                 \\
+        Diagonal   & 0.194                          & 0.314                    & 0.723                & 0.185                & 0.330                & 0.621                 \\
+        Tied Full  & {\color{red} 0.113 }           & {\color{red} 0.224 }     & 0.571                & {\color{red} 0.110 } & {\color{red} 0.207 } & 0.591                 \\
+        Tied Diag. & 0.161                          & 0.266                    & 0.591                & 0.152                & 0.262                & {\color{blue} 0.544 } \\
+
+        \hline
+        \multicolumn{7}{c}{RAW Features - PCA (m = 7)}                                                                                                                      \\
+        \hline
+        Full Cov.  & 0.174                          & 0.302                    & 0.708                & 0.161                & 0.301                & 0.664                 \\
+        Diagonal   & 0.220                          & 0.628                    & 0.833                & 0.217                & 0.626                & 0.678                 \\
+        Tied Full  & 0.153                          & 0.280                    & 0.612                & 0.147                & 0.268                & 0.622                 \\
+        Tied Diag. & 0.172                          & 0.306                    & 0.639                & 0.163                & 0.308                & 0.631                 \\
+
+        \hline
+        \multicolumn{7}{c}{RAW Features - PCA (m = 6)}                                                                                                                      \\
+        \hline
+        Full Cov.  & 0.172                          & 0.315                    & 0.662                & 0.157                & 0.309                & 0.626                 \\
+        Diagonal   & 0.220                          & 0.645                    & 0.818                & 0.215                & 0.648                & 0.687                 \\
+        Tied Full  & 0.156                          & 0.288                    & 0.602                & 0.147                & 0.268                & 0.607                 \\
+        Tied Diag. & 0.173                          & 0.310                    & 0.618                & 0.166                & 0.309                & 0.602                 \\
+
+        \hline
+        \multicolumn{7}{c}{Z-Normalized Features}                                                                                                                           \\
+        \hline
+        Full Cov.  & 0.141                          & 0.280                    & 0.680                & 0.140                & 0.283                & 0.646                 \\
+        Diagonal   & 0.194                          & 0.314                    & 0.723                & 0.185                & 0.330                & 0.621                 \\
+        Tied Full  & {\color{red} 0.113 }           & {\color{red} 0.224 }     & 0.571                & {\color{red} 0.110 } & {\color{red} 0.207 } & 0.591                 \\
+        Tied Diag. & 0.161                          & 0.267                    & 0.591                & 0.152                & 0.262                & {\color{blue} 0.544 } \\
+
+        \hline
+        \multicolumn{7}{c}{Gaussianized Features}                                                                                                                           \\
+        \hline
+        Full Cov.  & 0.154                          & 0.245                    & 0.717                & 0.151                & 0.241                & 0.622                 \\
+        Diagonal   & 0.152                          & 0.277                    & 0.602                & 0.154                & 0.285                & 0.574                 \\
+        Tied Full  & 0.132                          & {\color{blue} 0.237 }    & {\color{red} 0.532 } & 0.125                & 0.221                & {\color{red} 0.540 }  \\
+        Tied Diag. & 0.163                          & 0.292                    & 0.622                & 0.158                & 0.301                & 0.568                 \\
+    \end{tabular}
+    \captionof{table}{Experimental results - MVG classifiers}
+\end{center}
+
+The results are consistent with those obtained during validation, as they lead to the same conclusions:
+The MVG model with tied covariance remains the best performing one;
+PCA leads to a slight degradation in performance;
+Gaussianization obtains worse results than z-normalized features.
+
+\subsection{Logistic Regression}
+
+The effects of different values of the regularization parameter \(\lambda\) have been evaluated on the test set through the minimum DCF metric, and compared with the results obtained on validation data.
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-comparison-lr-z-normalized.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-comparison-lr-gaussianized.png}   \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\vspace*{-20pt}
+As for the validation set, regularization doesn't seem effective. Models have been re-trained with \(\lambda = 10^{-6}\). \\
+
+\renewcommand{\arraystretch}{1.2}
+\begin{tabular}{c|c@{\hspace{5pt}}c@{\hspace{5pt}}c|c@{\hspace{5pt}}c@{\hspace{5pt}}c}
+
+    Model                                        & \(\tilde{\pi}\)=0.5                & \(\tilde{\pi}\)=0.1          & \(\tilde{\pi}\)=0.9   & \(\tilde{\pi}\)=0.5  & \(\tilde{\pi}\)=0.1  & \(\tilde{\pi}\)=0.9   \\ [0.5ex]
+
+    \hline
+                                                 & \multicolumn{3}{c|}{Validation set} & \multicolumn{3}{|c}{Test set}                                                                                               \\
+    \hline
+    \multicolumn{7}{c}{RAW Features}                                                                                                                                                                               \\
+    \hline
+    Log Reg \((\lambda = 10^{-6}, \pi_T = 0.5)\) & {\color{blue} 0.116 }              & 0.219                        & {\color{blue} 0.530 } & {\color{red} 0.107 } & {\color{red} 0.199 } & 0.542                 \\
+    Log Reg \((\lambda = 10^{-6}, \pi_T = 0.1)\) & {\color{blue} 0.114 }              & {\color{red} 0.213 }         & 0.538                 & 0.110                & {\color{red} 0.199 } & 0.530                 \\
+    Log Reg \((\lambda = 10^{-6}, \pi_T = 0.9)\) & {\color{blue} 0.118 }              & 0.218                        & 0.534                 & 0.116                & 0.205                & 0.529                 \\
+
+    \hline
+    \multicolumn{7}{c}{Z-normalized Features}                                                                                                                                                                      \\
+    \hline
+    Log Reg \((\lambda = 10^{-6}, \pi_T = 0.5)\) & {\color{blue} 0.116 }              & 0.219                        & {\color{blue} 0.530 } & {\color{red} 0.107 } & {\color{red} 0.199 } & 0.542                 \\
+    Log Reg \((\lambda = 10^{-6}, \pi_T = 0.1)\) & {\color{red} 0.113 }               & {\color{red} 0.213 }         & 0.537                 & 0.110                & {\color{red} 0.199 } & 0.530                 \\
+    Log Reg \((\lambda = 10^{-6}, \pi_T = 0.9)\) & 0.120                              & 0.220                        & 0.537                 & 0.115                & 0.203                & 0.509                 \\
+
+    \hline
+    \multicolumn{7}{c}{Gaussianized Features}                                                                                                                                                                      \\
+    \hline
+    Log Reg \((\lambda = 10^{-6}, \pi_T = 0.5)\) & 0.126                              & 0.238                        & 0.534                 & 0.122                & 0.215                & {\color{blue} 0.485 } \\
+    Log Reg \((\lambda = 10^{-6}, \pi_T = 0.1)\) & 0.128                              & 0.229                        & 0.537                 & 0.125                & 0.211                & 0.504                 \\
+    Log Reg \((\lambda = 10^{-6}, \pi_T = 0.9)\) & 0.127                              & 0.238                        & {\color{red} 0.524 }  & 0.123                & 0.212                & {\color{red} 0.484 }  \\
+\end{tabular}
+\captionof{table}{Experimental results - Logistic Regression}
+
+The results are close to the ones obtained in the validation phase.
+The best minDCF values on the test set are obtained for \(\pi_T = 0.5 \) rather than with the proposed \(\pi_T = 0.1 \) but the difference in performance is very small.
+
+\subsection{Linear Support Vector Machines}
+
+Hyperparameter C has been re-tuned on the test set, values of minDCF have been plotted in the graphs below.
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-linear-svm-z-normalized-unbalanced.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-linear-svm-z-normalized-balanced.png}    \\
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-linear-svm-gaussianized-unbalanced.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-linear-svm-gaussianized-balanced.png}     \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+C $\approx 10^{-1}$ was selected for all applications and types of features during validation, the same value of C remains effective also on the test set.
+
+\subsection{Support Vector Machines with Polynomial kernel}
+
+A grid search of parameters C and c for the polynomial kernel SVM has been performed on the test set, comparisons with validation data are reported below.
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-25pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-polynomial-svm-z-normalized.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-polynomial-svm-gaussianized.png}   \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+Validation data demonstrated that C=1 and c=0.1 were good choices for the z-normalized features and C=1, c=1 for the Gaussianized features.
+Evaluation data shows that choosing $C=10^{-1}$ could have been a slightly better choice for both types of features. \\ \\
+
+{
+    \renewcommand{\arraystretch}{1.2}
+    \begin{tabular}{@{}c|c@{\hspace{5pt}}c@{\hspace{5pt}}c|c@{\hspace{5pt}}c@{\hspace{5pt}}c@{}}
+
+        Model                                  & \(\tilde{\pi}\)=0.5                & \(\tilde{\pi}\)=0.1          & \(\tilde{\pi}\)=0.9  & \(\tilde{\pi}\)=0.5  & \(\tilde{\pi}\)=0.1  & \(\tilde{\pi}\)=0.9  \\ [0.5ex]
+
+        \hline
+                                               & \multicolumn{3}{c|}{Validation set} & \multicolumn{3}{c}{Test set}                                                                                             \\
+        \hline
+
+        \multicolumn{7}{c}{Z-normalized Features}                                                                                                                                                              \\
+        \hline
+        Polynomial Kernel (c=0.1, $\pi_T=0.5$) & {\color{red} 0.103 }               & {\color{red} 0.220 }         & {\color{red} 0.525 } & {\color{red} 0.105 } & {\color{red} 0.199 } & {\color{red} 0.430 } \\
+        Polynomial Kernel (c=0.1, $\pi_T=0.1$) & 0.129                              & 0.228                        & 0.662                & 0.124                & 0.207                & 0.605                \\
+        Polynomial Kernel (c=0.1, $\pi_T=0.9$) & 0.141                              & 0.297                        & 0.594                & 0.176                & 0.570                & 0.503                \\
+
+        \hline
+        \multicolumn{7}{c}{Gaussianized Features}                                                                                                                                                              \\
+        \hline
+        Polynomial Kernel (c=1, $\pi_T=0.5$)   & 0.128                              & 0.255                        & 0.570                & 0.119                & 0.233                & 0.487                \\
+        Polynomial Kernel (c=1, $\pi_T=0.1$)   & 0.136                              & 0.244                        & 0.634                & 0.127                & 0.228                & 0.523                \\
+        Polynomial Kernel (c=1, $\pi_T=0.9$)   & 0.207                              & 0.485                        & 0.603                & 0.140                & 0.277                & 0.498                \\
+    \end{tabular}
+    \captionof{table}{Experimental results - Polynomial Kernel SVM}
+}
+
+\vspace{20pt}
+Experimental results for the polynomial kernel SVM showcase that c = 0.1 and $\pi_T = 0.5$ using z-normalized features is the best choice for all applications, as validation suggested.
+\clearpage
+
+\subsection{Support Vector Machines with RBF kernel}
+
+Evaluation of hyperparameter choices for the SVM with RBF kernel has also been performed through a grid search.
+
+\begin{figure}[H]
+    \begin{center}
+        \hspace*{-15pt}
+        \begin{tabular}{ccc}
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-rbf-svm-z-normalized.png} &
+            \includegraphics[width = 200pt]{img/evaluation_plots/evaluation-rbf-svm-gaussianized.png} &
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\vspace{-15pt}
+The RBF kernel has previously been used with C=$10^2$, $\gamma=0.1$ for both z-normalized and Gaussianized features.
+The hyperparameter tuning on the test set reveals that $\gamma=0.1$ was a good choice but perhaps C=10 would have led to slightly better results. \\
+
+\renewcommand{\arraystretch}{1.2}
+\begin{tabular}{@{}c|c@{\hspace{5pt}}c@{\hspace{5pt}}c|c@{\hspace{5pt}}c@{\hspace{5pt}}c@{}}
+
+    Model                                    & \(\tilde{\pi}\)=0.5                & \(\tilde{\pi}\)=0.1          & \(\tilde{\pi}\)=0.9  & \(\tilde{\pi}\)=0.5   & \(\tilde{\pi}\)=0.1   & \(\tilde{\pi}\)=0.9  \\ [0.5ex]
+
+    \hline
+                                            & \multicolumn{3}{c|}{Validation set} & \multicolumn{3}{c}{Test set}                                                                                               \\
+    \hline
+
+    \multicolumn{7}{c}{Z-normalized Features}                                                                                                                                                                  \\
+    \hline
+    RBF Kernel ($\gamma=0.1, \pi_T=0.5$) & {\color{blue} 0.115 }              & 0.236                        & 0.642                & {\color{blue} 0.118 } & 0.229                 & 0.595                \\
+    RBF Kernel ($\gamma=0.1, \pi_T=0.1$) & 0.123                              & {\color{blue} 0.211 }        & 0.830                & 0.122                 & {\color{blue} 0.209 } & 0.838                \\
+    RBF Kernel ($\gamma=0.1, \pi_T=0.9$) & 0.220                              & 0.690                        & 0.630                & 0.198                 & 0.716                 & 0.576                \\
+
+    \hline
+    \multicolumn{7}{c}{Gaussianized Features}                                                                                                                                                                  \\
+    \hline
+    RBF Kernel ($\gamma=0.1, \pi_T=0.5$) & {\color{red} 0.109 }               & 0.219                        & 0.631                & {\color{red} 0.111 }  & 0.231                 & 0.574                \\
+    RBF Kernel ($\gamma=0.1, \pi_T=0.1$) & 0.121                              & {\color{red} 0.206 }         & 0.773                & 0.119                 & {\color{red} 0.205 }  & 0.683                \\
+    RBF Kernel ($\gamma=0.1, \pi_T=0.9$) & 0.201                              & 0.460                        & {\color{red} 0.604 } & 0.176                 & 0.500                 & {\color{red} 0.545 } \\
+\end{tabular}
+\captionof{table}{Experimental results - RBF Kernel SVM}
+
+\vspace{15pt}
+The results for the RBF Kernel SVM are consistent with the ones obtained on the validation set. Using different priors is effective for different applications.
+
+\subsection{Gaussian Mixture Models}
+
+Different amounts of components have been evaluated over different kinds of GMM models for z-normalized and Gaussianized features, the results are reported in the following graphs.
+
+\begin{figure}[H]
+    \begin{center}
+        \begin{tabular}{ccc}
+            \hspace*{-65pt}
+            \includegraphics[width = 160pt]{img/evaluation_plots/evaluation-gmm-full-covariance.png}     &
+            \includegraphics[width = 160pt]{img/evaluation_plots/evaluation-gmm-diagonal-covariance.png} &
+            \includegraphics[width = 160pt]{img/evaluation_plots/evaluation-gmm-tied-covariance.png}       \\
+        \end{tabular}
+    \end{center}
+\end{figure}
+
+\vspace{-15pt}
+Z-normalized features produce the best results, as for the validation data.
+The optimal number of components remains 8 for the full covariance model and 32 for the tied covariance model.
+The most notable difference is the performance of the diagonal covariance model, which is better for 8 components rather than 32 as validation data suggested.
+
+\clearpage
+
+{
+    \hspace{-30pt}
+    \renewcommand{\arraystretch}{1.2}
+    \begin{tabular}{@{}c|c@{\hspace{5pt}}c@{\hspace{5pt}}c|c@{\hspace{5pt}}c@{\hspace{5pt}}c@{}}
+
+        Model                             & \(\tilde{\pi}\)=0.5                & \(\tilde{\pi}\)=0.1          & \(\tilde{\pi}\)=0.9  & \(\tilde{\pi}\)=0.5  & \(\tilde{\pi}\)=0.1  & \(\tilde{\pi}\)=0.9   \\
+
+        \hline
+                                        & \multicolumn{3}{c|}{Validation set} & \multicolumn{3}{c}{Test set}                                                                                              \\
+        \hline
+        \multicolumn{7}{c}{RAW Features}                                                                                                                                                                   \\
+        \hline
+        GMM Full Cov.
+8 components        & 0.130                              & 0.245                        & 0.630                & 0.143                & 0.264                & 0.606                 \\
+        GMM Tied Cov.
+32 components       & 0.147                              & 0.297                        & 0.715                & 0.148                & 0.295                & 0.601                 \\
+        GMM Diag.
+Cov.
+32 components      & 0.162                              & 0.288                        & 0.659                & 0.139                & 0.313                & 0.676                 \\
+        GMM Tied Diag.
+Cov.
+32 components & 0.159                              & 0.300                        & 0.691                & 0.155                & 0.306                & 0.644                 \\
+
+        \hline
+        \multicolumn{7}{c}{Z-normalized Features}                                                                                                                                                          \\
+        \hline
+        GMM Full Cov.
+8 components        & {\color{red} 0.118 }               & {\color{red} 0.224 }         & {\color{red} 0.562 } & {\color{red} 0.110 } & {\color{red} 0.218 } & {\color{blue} 0.586 } \\
+        GMM Tied Cov.
+32 components       & 0.129                              & 0.250                        & 0.615                & 0.147                & 0.279                & 0.592                 \\
+        GMM Diag.
+Cov.
+32 components      & 0.142                              & 0.276                        & 0.599                & 0.138                & 0.292                & 0.609                 \\
+        GMM Tied Diag.
+Cov.
+32 components & 0.153                              & 0.291                        & 0.647                & 0.149                & 0.296                & 0.645                 \\
+
+        \hline
+        \multicolumn{7}{c}{Gaussianized Features}                                                                                                                                                          \\
+        \hline
+        GMM Full Cov.
+8 components        & 0.128                              & 0.230                        & 0.680                & 0.135                & 0.249                & 0.590                 \\
+        GMM Tied Cov.
+32 components       & 0.139                              & {\color{blue} 0.226 }        & 0.668                & 0.153                & 0.253                & 0.603                 \\
+        GMM Diag.
+Cov.
+32 components      & 0.151                              & 0.284                        & 0.662                & 0.147                & 0.264                & {\color{red} 0.575 }  \\
+        GMM Tied Diag.
+Cov.
+32 components & 0.184                              & 0.293                        & 0.672                & 0.182                & 0.307                & 0.640                 \\
+    \end{tabular}
+    \captionof{table}{Experimental Results - GMM}
+}
+
+\vspace{20pt}
+Results obtained on the test set are in line with the ones observed during the validation phase.
+The Full Covariance GMM with 8 components is confirmed to be the best-performing model.
+
+\section{Conclusions}
+
+Different models have been evaluated for the pulsar classification task. Overall all classifiers are able to achieve good performances.
+PCA, LDA, and Gaussianization didn't provide substantial improvements while Z-normalization proved to be an effective preprocessing step.
+The similarity between experimental and validation results implies that evaluation and training populations behave similarly.
+The SVM with polynomial kernel was confirmed to be the best performing model by the experiments carried out on the test set.
+
 \bibliographystyle{unsrt}
 \bibliography{citations}
-
 \end{document}

### 2023-03-11 src/constants.py

@@ -15,8 +15,8 @@ LOG_REG_λ = [0, 1e-6, 1e-5, 1e-4, 1e-3, 1e-2, 1e-1, 1, 1e1]
 # Support Vector Machines
 
 SVM_C = [1e-3, 1e-2, 1e-1, 1e0, 1e1, 1e2]
-SVM_c = [1e-1, 1e-0, 1e1]
 SVM_γ = [1e-3, 1e-2, 1e-1]
+SVM_c = [1e-1, 1e-0, 1e1]
 SVM_K = 1.0
 
 # Gaussian Mixture Models

### 2023-03-11 src/data_processing/calibration.py

@@ -1,0 +1,54 @@
+from visualization.evaluation_plots import bayes_error_plot
+from models.train import logistic_regression_kfold
+from data_processing.validation import kfold
+from models.lr import LogisticRegression
+from constants import  K
+import numpy
+
+######################
+# Score Recalibation #
+######################
+
+
+def recalibrate_scores_kfold(scores: numpy.ndarray, labels: numpy.ndarray, π_tilde: float):
+    """Train LR models with K-fold to recalibrate scores"""
+
+    DTR_kfold, LTR_kfold, DVAL_kfold = numpy.empty(K, dtype=numpy.ndarray), numpy.empty(K, dtype=numpy.ndarray), numpy.empty(K, dtype=numpy.ndarray)
+    for i, (DTR, LTR, DVAL, _LVAL) in enumerate(kfold(scores, labels, K)):
+        DTR_kfold[i], LTR_kfold[i], DVAL_kfold[i] = DTR, LTR, DVAL
+    
+    # Train a LR model to recalibrate scores
+    recalibrated_scores = logistic_regression_kfold(
+        DTR_kfold, LTR_kfold, DVAL_kfold, labels, λ=0, πT=0.5, quadratic=False, application=(π_tilde, 1, 1), return_scores=True
+    )
+    
+    return recalibrated_scores - numpy.log(π_tilde / (1 - π_tilde))
+
+def recalibrate_scores(scores: numpy.ndarray, labels: numpy.ndarray, π_tilde: float):
+    """Train a LR model to recalibrate scores"""
+
+    lr = LogisticRegression(scores, labels, λ=0, πT=0.5, quadratic=False)
+    α, β = lr.w, lr.b
+    recalibrated_scores = α * scores + β - numpy.log(π_tilde / (1 - π_tilde))
+    
+    return recalibrated_scores.ravel(), α, β
+
+
+def recalibrate_models(models: dict, labels: numpy.ndarray, kfold: bool = False):
+    """Recalibrate models scores"""
+
+    recalibration_mode = recalibrate_scores_kfold if kfold else recalibrate_scores
+
+    for model, scores in models.items():
+        models[model] = recalibration_mode(numpy.array([scores]), labels, π_tilde=0.5)
+        print("{} recalibrated".format(model))
+        
+    return models
+
+
+def act_vs_min_DCF(models, LVAL):
+    """Draw Bayes error plots"""
+
+    for model, scores in models.items():
+
+        bayes_error_plot(model, scores, LVAL)

### 2023-03-11 src/data_processing/comparison.py

@@ -34,6 +34,7 @@ def optimal_bayes_decisions(scores: numpy.ndarray, π1: float, Cfn: float, Cfp: 
     """
     # C = numpy.array([[0, Cfn], [Cfp, 0]]) # Cost matrix
     # π = numpy.array([1 - π1, π1]) # Class priors
+    
     t = -numpy.log(π1 * Cfn / ((1 - π1) * Cfp))  # Threshold
     return (scores > t).astype(int)
 

### 2023-03-11 src/data_processing/data_load.py

@@ -58,7 +58,7 @@ def load_test() -> tuple[numpy.ndarray, numpy.ndarray]:
     return load_data("../data/Test.txt")
 
 
-def load_preprocessed(training_samples, training_labels, test_samples, test_labels):
+def load_preprocessed(training_samples, training_labels, test_samples):
 
     # Cached data files #
     folder = "../data/processed/"

### 2023-03-11 src/main.py

@@ -1,8 +1,21 @@
-from models.evaluation import PCA_m_selection, evaluate_MVG_models, evaluate_LR_models, LR_λ_selection, SVM_C_selection, SVM_C_kernel_selection, evaluate_SVM_models, GMM_components_selection, evaluate_GMM_models, act_vs_min_DCF, recalibrate_models
+from models.evaluation import evaluate_mvg, evaluate_lr, evaluate_svm, evaluate_gmm
 from data_processing.data_load import load_training, load_test, load_preprocessed
+from data_processing.calibration import act_vs_min_DCF, recalibrate_models
+from visualization.evaluation_plots import DET_plot
 from data_processing.utils import shuffle_data
 from models.train import train_best_models
 from models.svm import KernelType
+from models.validation import (
+    PCA_m_selection,
+    validate_MVG_models,
+    validate_LR_models,
+    LR_λ_selection,
+    SVM_C_selection,
+    SVM_C_kernel_selection,
+    validate_SVM_models,
+    GMM_components_selection,
+    validate_GMM_models,
+)
 
 if __name__ == "__main__":
 
@@ -28,8 +41,11 @@ if __name__ == "__main__":
         DVAL_kfold_z_normalized,
         LTR_kfold,
         LVAL_kfold,
-    ) = load_preprocessed(training_samples, training_labels, test_samples, test_labels)
+    ) = load_preprocessed(training_samples, training_labels, test_samples)
+
     LVAL = training_labels
+    LTR = training_labels
+    LTE = test_labels
 
     #################################
     # PCA hyper parameter selection #
@@ -41,8 +57,7 @@ if __name__ == "__main__":
     # MVG Models #
     ##############
 
-    # MVG MODELS
-    evaluate_MVG_models(
+    validate_MVG_models(
         DTR_kfold_raw,
         DTR_kfold_gaussianized,
         DTR_kfold_z_normalized,
@@ -50,7 +65,7 @@ if __name__ == "__main__":
         DVAL_kfold_raw,
         DVAL_kfold_gaussianized,
         DVAL_kfold_z_normalized,
-        LVAL
+        LVAL,
     )
 
     #######################
@@ -58,17 +73,12 @@ if __name__ == "__main__":
     #######################
 
     # Hyperparameter selection
[…]

### 2023-03-11 src/models/evaluation.py

@@ -1,402 +1,127 @@
-from constants import APPLICATIONS, K, LABELS, LOG_REG_λ, SVM_K, SVM_C, SVM_c, SVM_γ, GMM_STEPS
-from models.train import mvg_kfold, logistic_regression_kfold, svm_kfold, gmm_kfold
-from visualization.evaluation_plots import plot_minDCF, bayes_error_plot
+from models.train import train_mvg_models, train_log_reg_model, train_svm_model, train_gmm_models
 from data_processing.dimensionality_reduction import pca
 from visualization.print import print_separator
 from data_processing.utils import project_data
-from data_processing.validation import kfold
+from constants import APPLICATIONS, LABELS
 from models.svm import KernelType
 from models.mvg import MVGModel
 import numpy
 
 
-#########################
-# Hyperparameter tuning #
-#########################
-
-
-def PCA_m_selection(training_samples: numpy.ndarray, training_labels: numpy.ndarray) -> None:
-    """
-    Evaluate different values of the hyperparameter m for PCA
-
-    Args:
-        training_samples (numpy.ndarray): Training samples
-        training_labels (numpy.ndarray): Training labels for the samples
-    """
-
-    print_separator("PCA hyperparameter selection")
-
-    # Test m = 8, 7, 6, 5, 4, 3
-    from_components = 8
-    to_components = 3
-
-    for components in range(from_components, to_components - 1, -1):
-
-        # Divide training set into K folds
-        DTR_kfold_pca, DVAL_kfold_pca = numpy.empty(K, dtype=numpy.ndarray), numpy.empty(K, dtype=numpy.ndarray)
-        LTR_kfold = numpy.empty(K, dtype=numpy.ndarray)
-        for i, (DTR, LTR, DVAL, _LVAL) in enumerate(kfold(training_samples, training_labels, K)):
-            # Apply PCA projection
-            pca_projection_matrix = pca(DTR, components)
-            DTR_kfold_pca[i] = project_data(DTR, pca_projection_matrix)
-            DVAL_kfold_pca[i] = project_data(DVAL, pca_projection_matrix)
-            LTR_kfold[i] = LTR
-
-        for application in APPLICATIONS:
-
-            # Evaluate Gaussian models
[…]

### 2023-03-11 src/models/mvg.py

@@ -12,8 +12,8 @@ class MVGModel(Enum):
     TIED = "Tied Covariance"
     TIED_NAIVE = "Diagonal Tied Covariance"
 
+
 class MVG:
-    
     def __init__(self, type: MVGModel, DTR: numpy.ndarray, LTR: numpy.ndarray, labels: list[int]):
         """
         Train a multivariate gaussian model
@@ -44,7 +44,7 @@ class MVG:
             µ, Σ = self.train_tied_covariance_naive_bayes_mvg()
         else:
             raise ValueError("Invalid MVG model type")
-        
+
         self.µ, self.Σ = µ, Σ
 
     @staticmethod
@@ -157,7 +157,7 @@ class MVG:
 
         return µ, Σ
 
-    def score_samples(self, DTE: numpy.ndarray, priors: numpy.ndarray) -> numpy.ndarray:
+    def score_samples(self, DTE: numpy.ndarray, priors: numpy.ndarray) -> tuple[numpy.ndarray, numpy.ndarray]:
         """
         Compute the posterior probability for each class,
         Assign to each sample the class with the highest posterior probability
@@ -174,9 +174,13 @@ class MVG:
         # Compute class-conditional probabilities for each class
         S = [MVG.logpdf_GAU_ND(DTE, self.µ[c], self.Σ if self.is_tied else self.Σ[c]) for c in range(len(self.labels))]
 
+        # if α is not None and β is not None and π_tilde is not None:
+        # for i in range(len(S)):
+        #     S[i] = α * S[i] + β - numpy.log(π_tilde / (1 - π_tilde))
+
         # Compute Log Likelihood Ratios
         llr = S[1] - S[0]
-        
+
         # Combine the scores with prior information
         log_priors = vcol(numpy.log(priors))
         log_SJoint = S + log_priors

### 2023-03-11 src/models/train.py

@@ -1,34 +1,16 @@
+from data_processing.comparison import minimum_DCF, normalized_DCF, error_rate
+from data_processing.analytics import confusion_matrix
 from data_processing.utils import extended_data_matrix
-from data_processing.comparison import minimum_DCF
+from constants import LABELS, APPLICATIONS
 from models.lr import LogisticRegression
 from models.svm import SVM, KernelType
 from models.mvg import MVG, MVGModel
-from constants import LABELS
 from models.gmm import GMM
 from typing import Union
 import numpy
 
+
 # Single Model #
-
-def train_log_reg_model(
-    training_samples: numpy.ndarray,
-    training_labels: numpy.ndarray,
-    test_samples: numpy.ndarray,
-    test_labels: numpy.ndarray,
-    λ: float,
-    application: tuple[float, float, float],
-) -> float:
-
-    # Train model
-    model = LogisticRegression(training_samples, training_labels, λ, application[0])
-
-    # Compute scores
-    scores = model.score_samples(test_samples)
-
-    # Compute minimum DCF
-    minDCF = minimum_DCF(scores, test_labels, *application)
-
-    return minDCF
 
 
 def train_mvg_models(
@@ -36,31 +18,105 @@ def train_mvg_models(
     LTR: numpy.ndarray,
     DTE: numpy.ndarray,
     LTE: numpy.ndarray,
-    class_priors: list[float],
     labels: list[int],
     application: tuple[float, float, float],
+    models: list = MVGModel,
+    recalibration: tuple = None,
+    **kwargs,
 ) -> dict[MVGModel, float]:
 
-    minDCFs = {}
+    class_priors = [application[0], 1 - application[0]]
+    trained_models = {model_type: {} for model_type in models}
 
-    for model_type in MVGModel:
+    if recalibration is not None:
+        _, α, β = recalibration
+    else:
+        α, β = None, None
+
+    for model_type in models:
+
+        trained_models[model_type]["actDCF"] = {}
+        trained_models[model_type]["minDCF"] = {}
+        trained_models[model_type]["error_rate"] = {}
 
         # Train model
         model = MVG(model_type, DTR, LTR, labels)
 
-        # Compute scores for each class
-        model_scores = model.score_samples(DTE, class_priors)
+        for π_tilde, _, _ in APPLICATIONS:
 
[…]

### 2023-03-11 src/models/validation.py

@@ -1,0 +1,483 @@
+from models.train import mvg_kfold, logistic_regression_kfold, svm_kfold, gmm_kfold, train_log_reg_model, train_gmm_models, train_svm_model
+from constants import APPLICATIONS, K, LABELS, LOG_REG_λ, SVM_K, SVM_C, SVM_c, SVM_γ, GMM_STEPS
+from visualization.evaluation_plots import plot_minDCF
+from data_processing.dimensionality_reduction import pca
+from visualization.print import print_separator
+from data_processing.utils import project_data
+from data_processing.validation import kfold
+from models.svm import KernelType
+from models.mvg import MVGModel
+import numpy
+
+#########################
+# Hyperparameter tuning #
+#########################
+
+
+def PCA_m_selection(training_samples: numpy.ndarray, training_labels: numpy.ndarray) -> None:
+    """
+    Evaluate different values of the hyperparameter m for PCA
+
+    Args:
+        training_samples (numpy.ndarray): Training samples
+        training_labels (numpy.ndarray): Training labels for the samples
+    """
+
+    print_separator("PCA hyperparameter selection")
+
+    # Test components = 8, 7, 6, 5, 4, 3
+    from_components = 8
+    to_components = 3
+
+    for components in range(from_components, to_components - 1, -1):
+
+        # Divide training set into K folds
+        DTR_kfold_pca, DVAL_kfold_pca = numpy.empty(K, dtype=numpy.ndarray), numpy.empty(K, dtype=numpy.ndarray)
+        LTR_kfold = numpy.empty(K, dtype=numpy.ndarray)
+        for i, (DTR, LTR, DVAL, _LVAL) in enumerate(kfold(training_samples, training_labels, K)):
+            # Apply PCA projection
+            pca_projection_matrix = pca(DTR, components)
+            DTR_kfold_pca[i] = project_data(DTR, pca_projection_matrix)
+            DVAL_kfold_pca[i] = project_data(DVAL, pca_projection_matrix)
+            LTR_kfold[i] = LTR
+
+        for application in APPLICATIONS:
+
+            # K-fold cross validation of Gaussian models
+            validated_mvg_models = mvg_kfold(DTR_kfold_pca, LTR_kfold, DVAL_kfold_pca, training_labels, LABELS, application)
+            print("# PCA m = {} - {} #".format(components, application))
[…]

### 2023-03-11 src/visualization/evaluation_plots.py

@@ -1,5 +1,5 @@
-from data_processing.comparison import optimal_bayes_decisions, normalized_DCF, minimum_DCF
-from data_processing.analytics import confusion_matrix
+from data_processing.comparison import optimal_bayes_decisions, optimal_bayes_decisions_threshold, normalized_DCF, minimum_DCF
+from data_processing.analytics import confusion_matrix, fnr_fpr
 import matplotlib.pyplot as plt
 import numpy
 
@@ -35,16 +35,21 @@ def bayes_error_plot(title: str, scores: numpy.ndarray, labels: numpy.ndarray) -
     plt.show()
 
 
-def plot_minDCF(title: str, hyperparameter: str, x_axis: list, applications: dict, log_scale: bool = False, max_y: float = 1.1, param_name='$\tilde{\pi}$', extra = '0.5) - ') -> None:
+def plot_minDCF(title: str, hyperparameter: str, x_axis: list, applications: dict, log_scale: bool = False, max_y: float = 1.1, param_name=r'$\tilde{\pi}$', extra = '0.5) - ', comparison_applications: None = dict) -> None:
 
     colors = ["r", "b", "g"]
     xi = x_axis if log_scale else list(range(len(x_axis)))
+    application_type = " [Eval]" if comparison_applications is not None else ""
 
     for i, (application, minDCFs) in enumerate(applications.items()):
 
-        param_value = str(application[0]) + ')' if type(application) is tuple else extra + str(application)
-        
+        param_value = (str(application[0]) + ')' if type(application) is tuple else extra + str(application)) + application_type
         plt.plot(xi, minDCFs, label=r'minDCF(' + param_name + ' = ' + param_value, color=colors[i])
+
+    if comparison_applications is not None:
+        for i, (application, minDCFs) in enumerate(comparison_applications.items()):
+            param_value = (str(application[0]) + ')' if type(application) is tuple else extra + str(application)) + ' [Val]'
+            plt.plot(xi, minDCFs, label=r'minDCF(' + param_name + ' = ' + param_value, color=colors[i], linestyle='dashed')
 
     plt.title(title)
     plt.xlabel(hyperparameter)
@@ -54,5 +59,39 @@ def plot_minDCF(title: str, hyperparameter: str, x_axis: list, applications: dic
     else:
         plt.xticks(xi, x_axis)
     plt.ylim([0, max_y])
+    plt.legend()
+    plt.show()
+    
+    
[…]

### 2023-03-11 src/visualization/print.py

@@ -8,7 +8,7 @@ def print_separator(title: str = None):
         title (str, optional): Optional title to display in the middle of the separator. Defaults to None.
     """
     
-    SEPARATOR_LENGTH = 40
+    SEPARATOR_LENGTH = 60
     SEPARATOR = "#"
     print("\n" + SEPARATOR * SEPARATOR_LENGTH)
     
